---
title: "Getting Started with OMOPPregnancy"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with OMOPPregnancy}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

```{r setup}
library(OMOPPregnancy)
library(DatabaseConnector)
library(dplyr)
```

## Introduction

The OMOPPregnancy package implements the HIPPS (Hierarchy and rule-based pregnancy episode Inference integrated with Pregnancy Progression Signatures) algorithm for identifying pregnancy episodes in OMOP CDM databases. This vignette demonstrates how to use the package with your OMOP CDM database.

## Prerequisites

Before using the package, ensure you have:

1. Access to an OMOP CDM v5.3+ database
2. Required R packages installed
3. Database connection credentials
4. Sufficient permissions to create temporary tables

## Basic Usage

### Step 1: Create Database Connection

For a PostgreSQL database:

```{r}
connectionDetails <- createConnectionDetails(
  dbms = "postgresql",
  server = "localhost/mydb",
  user = "username",
  password = "password",
  port = 5432
)
```

For Microsoft SQL Server:

```{r}
connectionDetails <- createConnectionDetails(
  dbms = "sql server",
  server = "myserver.database.windows.net",
  user = "username",
  password = "password",
  port = 1433
)
```

For BigQuery (Google Cloud):

```{r}
connectionDetails <- createConnectionDetails(
  dbms = "bigquery",
  connectionString = "jdbc:bigquery://https://www.googleapis.com/bigquery/v2:443",
  user = "service-account@project.iam.gserviceaccount.com",
  password = "path/to/keyfile.json"
)
```

### Step 2: Run the HIPPS Algorithm

Basic execution with default settings:

```{r}
results <- run_hipps(
  connectionDetails = connectionDetails,
  cdmDatabaseSchema = "cdm_schema",
  outputFolder = "output/"
)
```

With custom configuration:

```{r}
# Load custom configuration
config <- load_config(mode = "generic")

# Modify settings
config$algorithm$hip$min_age <- 18  # Change minimum age
config$algorithm$hip$max_age <- 50  # Change maximum age

# Run with custom config
results <- run_hipps(
  connectionDetails = connectionDetails,
  cdmDatabaseSchema = "cdm_schema",
  resultsDatabaseSchema = "scratch_schema",
  outputFolder = "output/",
  config = config
)
```

### Step 3: Examine Results

View summary statistics:

```{r}
# Number of episodes identified
results$summary$n_episodes

# Number of unique persons with pregnancies
results$summary$n_persons

# Distribution by outcome type
results$summary$episodes_by_outcome
```

Access the full episode data:

```{r}
# Get local copy of episodes
episodes <- collect(results$episodes)

# View first few episodes
head(episodes)

# Summarize by year
episodes %>%
  mutate(year = lubridate::year(episode_start_date)) %>%
  group_by(year, category) %>%
  summarise(n = n()) %>%
  print()
```

## Using with All of Us

If you're working within the All of Us Research Workbench:

```{r}
# Use All of Us mode
results <- run_hipps(
  mode = "allofus",
  cdmDatabaseSchema = "aou_cdm",
  outputFolder = "output/"
)
```

## Comparing Across Sites

To compare results between different OMOP CDM databases:

```{r}
# Run on Site A
site_a_results <- run_hipps(
  connectionDetails = site_a_connection,
  cdmDatabaseSchema = "cdm",
  outputFolder = "output/site_a/"
)

# Run on Site B
site_b_results <- run_hipps(
  connectionDetails = site_b_connection,
  cdmDatabaseSchema = "cdm",
  outputFolder = "output/site_b/"
)

# Compare results
comparison <- compare_results(
  reference = site_a_results,
  comparison = site_b_results,
  stratify_by = c("age_group", "year", "outcome_type")
)

# Generate comparison report
generate_report(
  comparison,
  output_file = "comparison_report.html"
)
```

## Configuration Options

### Algorithm Parameters

Customize the algorithm behavior:

```{r}
config <- list(
  algorithm = list(
    hip = list(
      min_age = 15,        # Minimum age for inclusion
      max_age = 55,        # Maximum age for inclusion
      outcome_hierarchy = c("LB", "SB", "ECT", "AB", "SA", "DELIV")
    ),
    pps = list(
      min_records = 2,     # Minimum records for PPS episode
      max_gap_months = 10  # Maximum gap between concepts
    )
  )
)
```

### Database Settings

Optimize for your database platform:

```{r}
config$database <- list(
  batch_size = 10000,      # Records per batch
  use_temp_tables = TRUE,  # Use temporary tables
  parallel = FALSE         # Enable parallel processing
)
```

## Troubleshooting

### Connection Issues

If you encounter connection problems:

```{r}
# Test connection
con <- create_connection(connectionDetails, cdmDatabaseSchema = "cdm")
validate_connection(con, "cdm")
```

### Memory Issues

For large databases, use batching:

```{r}
config$database$batch_size <- 5000  # Smaller batch size
config$output$save_intermediate <- TRUE  # Save intermediate results
```

### Missing Concepts

Check concept availability:

```{r}
# Load concept sets
concepts <- load_concept_sets()

# Check if your database has required concepts
con <- create_connection(connectionDetails)
concept_tbl <- get_cdm_table(con, "concept", "cdm")

available_concepts <- concept_tbl %>%
  filter(concept_id %in% !!concepts$HIP_concepts$concept_id) %>%
  count() %>%
  pull(n)

message(paste("Found", available_concepts, "of", 
              nrow(concepts$HIP_concepts), "HIP concepts"))
```

## Advanced Usage

### Custom Concept Sets

Use your own concept sets:

```{r}
# Load custom concepts
my_concepts <- read.csv("my_hip_concepts.csv")

# Save to package format
saveRDS(my_concepts, "custom_concepts/HIP_concepts.rds")

# Run with custom concepts
results <- run_hipps(
  connectionDetails = connectionDetails,
  cdmDatabaseSchema = "cdm",
  conceptDir = "custom_concepts/",
  outputFolder = "output/"
)
```

### Parallel Processing

For large datasets, enable parallel processing:

```{r}
# Enable parallel execution
config$database$parallel <- TRUE

# Set number of cores
options(mc.cores = 4)

# Run with parallel processing
results <- run_hipps(
  connectionDetails = connectionDetails,
  cdmDatabaseSchema = "cdm",
  config = config,
  outputFolder = "output/"
)
```

## Output Files

The package generates several output files:

- `pregnancy_episodes.rds`: R data file with all episodes
- `pregnancy_episodes.csv`: CSV file for sharing
- `summary_statistics.txt`: Summary of results
- `execution_log.txt`: Detailed execution log

## Next Steps

- See the [Algorithm Reference](../docs/ALGORITHM_REFERENCE.md) for detailed algorithm documentation
- See the [Comparison Guide](../docs/COMPARISON_GUIDE.md) for multi-site studies
- Report issues at [GitHub Issues](https://github.com/yourusername/OMOPPregnancy/issues)