---
title: "Pregnancy Algorithm Validation Report"
subtitle: "Comparison of Algorithm Predictions vs Obstetric Registry Ground Truth"
author: "OMOP Pregnancy Team"
date: "`r Sys.Date()`"
format:
  html:
    toc: true
    toc-depth: 3
    toc-location: left
    code-fold: true
    code-summary: "Show code"
    theme: flatly
    self-contained: true
execute:
  echo: true
  warning: false
  message: false
params:
  # How to reach the database: "jdbc" (DatabaseConnector via env_file) or
  # "spark" (the Databricks notebook's Spark session; CDM = omop_database.omop_schema)
  connection_type: "jdbc"
  # Pipeline output to evaluate (path relative to this file), or "latest" for
  # the newest pregnancy_episodes_*.csv / pregnancy_analysis_*.csv in ../output
  prediction_file: "../output/pregnancy_episodes_2026-03-08.csv"
  # Episode matching window (days) around ground-truth end dates
  date_window_days: 30
  # Prediction filters: deliveries only (>= 20 weeks) and starts before this date
  min_ga_days: 140
  max_start_date: "2025-01-01"
  # Ground-truth source: "pmap_sqlserver" (PMAP OB registry tables on the SQL
  # Server instance) or "edw_databricks" (EDW pregnancy/birth fact tables on
  # Databricks). Both produce the same episode frame for the analyses below.
  gt_source: "pmap_sqlserver"
  # pmap_sqlserver: registry and linkage databases
  gt_scratch_database: "Obstetrics_Minhas_IRB00501137_Scratch"
  gt_projection_database: "Obstetrics_Minhas_IRB00501137_Projection"
  # edw_databricks: catalog holding the EDW fact tables and the registry id map
  edw_catalog: "obstetrics_irb00501137"
  edw_phi_schema: "phi"
  edw_idmap_schema: "omop"
  # OMOP CDM used by the complication section: database (or Databricks catalog)
  # and schema of the CDM tables, e.g. "Obstetrics_Minhas_IRB00501137_OMOP"/"dbo"
  # on SQL Server or "omop"/"data" on Databricks
  omop_database: "Obstetrics_Minhas_IRB00501137_OMOP"
  omop_schema: "dbo"
  # .env with the database connection (project root by default)
  env_file: "../.env"
---

```{r setup}
#| include: false

# Load required packages
library(dplyr)
library(tidyr)
library(purrr)
library(lubridate)
library(ggplot2)
library(scales)
library(knitr)
library(kableExtra)
library(janitor)

# Source the connection functions from the project (paths relative to this file)
source("../R/00_connection/create_connection.R")
source("../R/00_connection/db_backend.R")

# One place that opens the database connection for either backend
open_report_connection <- function() {
  if (identical(params$connection_type, "spark")) {
    create_spark_connection(cdm_schema = paste(params$omop_database, params$omop_schema, sep = "."))
  } else {
    create_connection_from_env(params$env_file)
  }
}

# Database names used in the ground-truth queries
gt_source        <- match.arg(params$gt_source, c("pmap_sqlserver", "edw_databricks"))
gt_scratch_db    <- params$gt_scratch_database
gt_projection_db <- params$gt_projection_database
edw_catalog      <- params$edw_catalog
edw_phi_schema   <- params$edw_phi_schema
edw_idmap_schema <- params$edw_idmap_schema
omop_db          <- params$omop_database
omop_schema_name <- params$omop_schema
cat("Ground-truth source:", gt_source, "\n")

# Set ggplot theme
theme_set(theme_minimal(base_size = 12))
```

# Executive Summary

This report validates the pregnancy identification algorithm by comparing its predictions against ground truth data from the PMAP OB registry. The validation covers:

- **Episode-level matching**: Matching predicted episodes to ground truth using a ±`r params$date_window_days` day window
- **Outcome classification**: Accuracy of outcome category predictions (LB, SB, ECT, AB, SA, DELIV, PREG)
- **Gestational age estimation**: Comparison of predicted vs ground truth gestational age
- **Person-level aggregation**: Total pregnancy counts per person

```{r load-predictions}

# Load algorithm predictions
load_algorithm_predictions <- function(file_path) {
  stopifnot(is.character(file_path), length(file_path) == 1L, nzchar(file_path))
  if (!file.exists(file_path)) stop("File not found: ", file_path)

  ext <- tolower(tools::file_ext(file_path))

  predictions <- if (ext == "rds") {
    readRDS(file_path)
  } else if (ext %in% c("csv", "txt")) {
    read.csv(file_path, stringsAsFactors = FALSE, check.names = FALSE)
  } else {
    stop("Unsupported file extension: .", ext, " (supported: rds, csv, txt)")
  }

  # Require deps explicitly (fail fast with clear message)
  if (!requireNamespace("dplyr", quietly = TRUE)) stop("Package 'dplyr' is required.")
  if (!requireNamespace("janitor", quietly = TRUE)) stop("Package 'janitor' is required.")

  # Helper: robust Date coercion (supports Date/POSIXct/character/numeric)
  to_date_safe <- function(x) {
    if (inherits(x, "Date")) return(x)
    if (inherits(x, "POSIXt")) return(as.Date(x))
    if (is.numeric(x)) return(as.Date(x, origin = "1970-01-01"))
    suppressWarnings(as.Date(x))
  }

  predictions <- predictions %>%
    janitor::clean_names()

  # If *_x columns exist (merge artifacts), coalesce into canonical fields
  if (!"episode_start_date" %in% names(predictions) && "episode_start_date_x" %in% names(predictions)) {
    predictions$episode_start_date <- predictions$episode_start_date_x
  }
  if (!"episode_end_date" %in% names(predictions) && "episode_end_date_x" %in% names(predictions)) {
    predictions$episode_end_date <- predictions$episode_end_date_x
  }

  # Validate required columns
  required_cols <- c(
    "person_id", "episode_number", "episode_start_date", "episode_end_date",
    "outcome_category", "gestational_age_days", "algorithm_used", "precision_category"
  )
  missing <- setdiff(required_cols, names(predictions))
  if (length(missing) > 0) {
    stop("Missing required columns after clean_names(): ", paste(missing, collapse = ", "))
  }

  # Clean / type-cast and select canonical fields
  predictions <- predictions %>%
    dplyr::mutate(
      person_id = suppressWarnings(as.integer(person_id)),
      episode_number = suppressWarnings(as.integer(episode_number)),
      episode_start_date = to_date_safe(episode_start_date),
      episode_end_date = to_date_safe(episode_end_date),
      gestational_age_days = suppressWarnings(as.integer(gestational_age_days)),
      outcome_category = as.character(outcome_category),
      algorithm_used = as.character(algorithm_used),
      precision_category = as.character(precision_category)
    ) %>%
    dplyr::select(dplyr::all_of(required_cols)) %>%
    dplyr::distinct() %>%
    dplyr::filter(episode_start_date < as.Date(params$max_start_date))

  return(predictions)
}

# Load predictions ("latest" = newest pregnancy_episodes_*.csv in ../output)
prediction_file <- params$prediction_file
if (identical(prediction_file, "latest")) {
  candidates <- list.files("../output", pattern = "^pregnancy_(episodes|analysis)_.*\\.csv$", full.names = TRUE)
  if (length(candidates) == 0) stop("No pregnancy_episodes_*.csv or pregnancy_analysis_*.csv found in ../output")
  prediction_file <- candidates[which.max(file.info(candidates)$mtime)]
}
cat("Prediction file:", prediction_file, "\n")
algo_predictions <- load_algorithm_predictions(prediction_file)
algo_predictions <- algo_predictions %>%
  dplyr::filter(gestational_age_days >= params$min_ga_days)
cat(
  "Loaded", nrow(algo_predictions), "algorithm predictions for",
  dplyr::n_distinct(algo_predictions$person_id), "unique persons\n"
)
```

```{r db-connect1}
# Connect to database using project connection function
con <- open_report_connection()
if (is.null(con)) stop("Connection returned NULL.")
if (!db_is_valid(con)) stop("Database connection is not valid.")

cat("Database connection established\n")

```

```{r query-ground-truth}
#| eval: !expr gt_source == "pmap_sqlserver"

# Query ground truth tables

# 1. Pregnancy episodes
pregnancies_sql <- paste0("
SELECT
  EMRN,
  PregnancyEstimatedStartDate,
  PregnancyEstimatedEndDate,
  PregnancyGravidaCount,
  PregnancyParaCount,
  PregnancyAbortionCount,
  PregnancyTherapeuticAbortionCount,
  PregnancySpontaneousAbortionCount,
  PregnancyEctopicCount
FROM ", gt_scratch_db, ".phi.PMAP_OB_PREGNANCIES
WHERE IsHistorical = 0 AND PregnancyEstimatedStartDate < '", params$max_start_date, "';
")

gt_pregnancies <- db_query(con, pregnancies_sql) %>%
  clean_names() %>%
  mutate(
    pregnancy_estimated_start_date = as.Date(pregnancy_estimated_start_date),
    pregnancy_estimated_end_date = as.Date(pregnancy_estimated_end_date)
  )

cat("Loaded", nrow(gt_pregnancies), "ground truth pregnancy episodes\n")
cat("Sample EMRNs from pregnancies:", head(gt_pregnancies$emrn, 5), "\n")

# 2. Delivery outcomes
deliveries_sql <- paste0("
SELECT
  mom_emrn AS emrn,
  GestationalAgeDays,
  LivingStatus,
  BornAlive,
  NeonatalDemise
FROM ", gt_scratch_db, ".phi.PMAP_OB_DEL
")

gt_deliveries <- db_query(con, deliveries_sql) %>%
  clean_names()

cat("Loaded", nrow(gt_deliveries), "ground truth delivery records\n")

# 3. EMRN to cohort_id linkage
emrn_cohort_sql <- paste0("
SELECT emrn, cohort_id
FROM ", gt_projection_db, ".phi.derived_epic_patient
")

emrn_to_cohort <- db_query(con, emrn_cohort_sql) %>%
  clean_names()

cat("Loaded", nrow(emrn_to_cohort), "EMRN to cohort_id mappings\n")
cat("Sample EMRNs from linkage:", head(emrn_to_cohort$emrn, 5), "\n")

# Check EMRN overlap between pregnancies and linkage
preg_emrns <- unique(gt_pregnancies$emrn)
link_emrns <- unique(emrn_to_cohort$emrn)
emrn_overlap <- length(intersect(preg_emrns, link_emrns))
cat("EMRN overlap (pregnancies ∩ linkage):", emrn_overlap, "/", length(preg_emrns), "\n")

# 4. cohort_id (pmap_id) to person_id linkage
cohort_person_sql <- paste0("
SELECT pmap_id AS cohort_id, person_id
FROM ", omop_db, ".", omop_schema_name, ".registry_idmap
")

cohort_to_person <- db_query(con, cohort_person_sql) %>%
  clean_names()

cat("Loaded", nrow(cohort_to_person), "cohort_id to person_id mappings\n")
```

```{r id-linkage}
#| eval: !expr gt_source == "pmap_sqlserver"

# Build complete ID linkage: EMRN -> cohort_id -> person_id
build_id_linkage <- function(emrn_cohort, cohort_person) {
  linkage <- emrn_cohort %>%
    inner_join(cohort_person, by = "cohort_id") %>%
    select(emrn, person_id) %>%
    distinct()

  return(linkage)
}

# Debug: Check cohort_id overlap before building linkage
cohort_ids_from_emrn <- unique(emrn_to_cohort$cohort_id)
cohort_ids_from_person <- unique(cohort_to_person$cohort_id)
cohort_overlap <- length(intersect(cohort_ids_from_emrn, cohort_ids_from_person))
cat("Cohort_id overlap (emrn_table ∩ person_table):", cohort_overlap, "\n")
cat("Sample cohort_ids from emrn table:", head(cohort_ids_from_emrn, 5), "\n")
cat("Sample cohort_ids from person table:", head(cohort_ids_from_person, 5), "\n")

id_linkage <- build_id_linkage(emrn_to_cohort, cohort_to_person)

cat("\nCreated ID linkage table with", nrow(id_linkage), "unique EMRN-person_id pairs\n")
cat("Columns in id_linkage:", names(id_linkage), "\n")

# Debug: Show sample of linkage
cat("Sample person_ids in linkage:", head(id_linkage$person_id, 10), "\n")
cat("Sample emrns in linkage:", head(id_linkage$emrn, 5), "\n")

# Check linkage coverage
algo_persons <- n_distinct(algo_predictions$person_id)
algo_person_ids <- unique(algo_predictions$person_id)
linked_persons <- sum(algo_person_ids %in% id_linkage$person_id)
cat("Algorithm persons with ID linkage:", linked_persons, "/", algo_persons,
    "(", round(100 * linked_persons / algo_persons, 1), "%)\n")

# Debug: Show sample algorithm person_ids
cat("Sample algorithm person_ids:", head(algo_person_ids, 10), "\n")

# Check for any overlap
overlap_count <- length(intersect(algo_person_ids, id_linkage$person_id))
cat("Overlapping person_ids between algorithm and linkage:", overlap_count, "\n")
```

```{r derive-outcomes}
#| eval: !expr gt_source == "pmap_sqlserver"

# Derive outcome category from ground truth cumulative counts
derive_ground_truth_outcome <- function(pregnancies_df, deliveries_df, id_linkage) {

  # Ensure count columns are numeric (may come as character from SQL)
  pregnancies_df <- pregnancies_df %>%
    mutate(
      pregnancy_gravida_count = as.numeric(pregnancy_gravida_count),
      pregnancy_para_count = as.numeric(pregnancy_para_count),
      pregnancy_abortion_count = as.numeric(pregnancy_abortion_count),
      pregnancy_therapeutic_abortion_count = as.numeric(pregnancy_therapeutic_abortion_count),
      pregnancy_spontaneous_abortion_count = as.numeric(pregnancy_spontaneous_abortion_count),
      pregnancy_ectopic_count = as.numeric(pregnancy_ectopic_count)
    )

  # Join pregnancies with ID linkage to get person_id
  pregnancies_linked <- pregnancies_df %>%
    inner_join(id_linkage, by = "emrn")

  # Ensure delivery columns are proper types
  deliveries_df <- deliveries_df %>%
    mutate(
      gestational_age_days = as.numeric(gestational_age_days),
      born_alive = as.numeric(born_alive)
    )

  # Join with deliveries (if available) for BornAlive info
  # Deliveries may have multiple records per EMRN, take the one closest to pregnancy end
  deliveries_by_emrn <- deliveries_df %>%
    group_by(emrn) %>%
    summarise(
      gestational_age_days = first(na.omit(gestational_age_days)),
      born_alive = first(na.omit(born_alive)),
      living_status = first(na.omit(living_status)),
      .groups = "drop"
    )

  # Combine and derive outcomes
  derived <- pregnancies_linked %>%
    left_join(deliveries_by_emrn, by = "emrn") %>%
    group_by(person_id) %>%
    arrange(pregnancy_gravida_count) %>%
    mutate(
      # Calculate deltas from previous episode
      prev_para = lag(pregnancy_para_count, default = 0),
      prev_ab = lag(pregnancy_abortion_count, default = 0),
      prev_therapeutic = lag(pregnancy_therapeutic_abortion_count, default = 0),
      prev_spontaneous = lag(pregnancy_spontaneous_abortion_count, default = 0),
      prev_ectopic = lag(pregnancy_ectopic_count, default = 0),

      delta_para = pregnancy_para_count - prev_para,
      delta_ab = pregnancy_abortion_count - prev_ab,
      delta_therapeutic = pregnancy_therapeutic_abortion_count - prev_therapeutic,
      delta_spontaneous = pregnancy_spontaneous_abortion_count - prev_spontaneous,
      delta_ectopic = pregnancy_ectopic_count - prev_ectopic,

      # Derive outcome category following Matcho hierarchy
      gt_outcome_category = case_when(
        # Live birth: para increased and born alive
        delta_para > 0 & !is.na(born_alive) & born_alive == 1 ~ "LB",
        # Stillbirth: para increased and not born alive (or living_status indicates)
        delta_para > 0 & (!is.na(born_alive) & born_alive == 0) ~ "SB",
        delta_para > 0 & (!is.na(living_status) & tolower(living_status) %in% c("stillborn", "fetal demise")) ~ "SB",
        # Ectopic pregnancy
        delta_ectopic > 0 ~ "ECT",
        # Therapeutic/Induced abortion
        delta_therapeutic > 0 ~ "AB",
        # Spontaneous abortion
        delta_spontaneous > 0 ~ "SA",
        # Generic abortion (when not specified as therapeutic or spontaneous)
        delta_ab > 0 & delta_therapeutic == 0 & delta_spontaneous == 0 ~ "AB",
        # Delivery without specific birth info
        delta_para > 0 ~ "DELIV",
        # Pregnancy without recorded outcome
        TRUE ~ "PREG"
      )
    ) %>%
    ungroup() %>%
    select(
      person_id,
      emrn,
      gt_episode_num = pregnancy_gravida_count,
      gt_start = pregnancy_estimated_start_date,
      gt_end = pregnancy_estimated_end_date,
      gt_outcome_category,
      gt_ga_days = gestational_age_days,
      born_alive,
      living_status
    )

  return(derived)
}

gt_episodes <- derive_ground_truth_outcome(gt_pregnancies, gt_deliveries, id_linkage)

cat("Derived outcomes for", nrow(gt_episodes), "ground truth episodes\n")
cat("Unique persons in ground truth:", n_distinct(gt_episodes$person_id), "\n")

# Debug: Check overlap between algorithm and ground truth persons
algo_persons_set <- unique(algo_predictions$person_id)
gt_persons_set <- unique(gt_episodes$person_id)
person_overlap <- length(intersect(algo_persons_set, gt_persons_set))
cat("Person overlap (algo ∩ GT):", person_overlap, "\n")

# Debug: Sample dates
cat("\nSample algorithm end dates:", head(algo_predictions$episode_end_date, 5), "\n")
cat("Sample GT end dates:", head(gt_episodes$gt_end, 5), "\n")

cat("\nGround truth outcome distribution:\n")
gt_episodes %>%
  count(gt_outcome_category) %>%
  arrange(desc(n)) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)

```

# Episode-level matching
```{r edw-ground-truth}
#| eval: !expr gt_source == "edw_databricks"

# =========================================================
# Ground truth from the EDW obstetric fact tables (Databricks)
#   <catalog>.<phi schema>.edw_pregnancyfact   one row per pregnancy
#   <catalog>.<phi schema>.edw_birthfact       one row per baby (PregnancyKey)
#   <catalog>.<idmap schema>.registry_idmap    PMAP_id -> OMOP person_id
# Produces gt_episodes with the same columns as the PMAP adapter:
#   person_id, gt_episode_num, gt_start, gt_end, gt_outcome_category, gt_ga_days
# =========================================================

edw_table <- function(schema, table) paste0(edw_catalog, ".", schema, ".", table)

edw_sql <- paste0("
WITH births AS (
  SELECT
    PregnancyKey,
    COUNT(*)                                                     AS n_births,
    SUM(CASE WHEN BornAlive = 1 THEN 1 ELSE 0 END)               AS n_born_alive,
    SUM(CASE WHEN LivingStatus = 'Fetal Demise' THEN 1 ELSE 0 END) AS n_fetal_demise,
    MAX(GestationalAgeDays)                                      AS ga_days_birth
  FROM ", edw_table(edw_phi_schema, "edw_birthfact"), "
  GROUP BY PregnancyKey
)
SELECT
  r.person_id                     AS person_id,
  p.cohort_id                     AS cohort_id,
  p.PregnancyKey                  AS pregnancy_key,
  p.PregnancyOutcome              AS pregnancy_outcome,
  p.HasDelivery                   AS has_delivery,
  p.HadFetalDemise                AS had_fetal_demise,
  p.HadNeonatalDemise             AS had_neonatal_demise,
  p.PregnancyEstimatedStartDate   AS pregnancy_estimated_start_date,
  p.EpisodeStartDate              AS episode_start_date,
  p.PregnancyEstimatedEndDate     AS pregnancy_estimated_end_date,
  p.EpisodeEndDate                AS episode_end_date,
  p.LastDeliveryDate              AS last_delivery_date,
  p.LastDeliveryGestationalAge    AS last_delivery_gestational_age,
  b.n_births                      AS n_births,
  b.n_born_alive                  AS n_born_alive,
  b.n_fetal_demise                AS n_fetal_demise,
  b.ga_days_birth                 AS ga_days_birth
FROM ", edw_table(edw_phi_schema, "edw_pregnancyfact"), " p
INNER JOIN ", edw_table(edw_idmap_schema, "registry_idmap"), " r
  ON p.cohort_id = r.PMAP_id
LEFT JOIN births b
  ON p.PregnancyKey = b.PregnancyKey
WHERE p.PregnancyOutcome IS NULL OR p.PregnancyOutcome <> '*Deleted'
")

# Columns are aliased to snake_case in the SQL; clean_names() only normalizes case
edw_raw <- db_query(con, edw_sql) %>%
  clean_names()

cat("Loaded", nrow(edw_raw), "EDW pregnancy episodes for",
    n_distinct(edw_raw$person_id), "linked persons\n")

to_date_any <- function(x) {
  if (inherits(x, "Date")) return(x)
  if (inherits(x, "POSIXt")) return(as.Date(x))
  suppressWarnings(as.Date(as.character(x)))
}

# Outcome mapping to the algorithm's categories (Matcho hierarchy):
#   Term / Preterm with any liveborn (or no birth rows and no fetal demise) -> LB
#   Term / Preterm where every birth was a fetal demise (or HadFetalDemise
#     with no birth rows)                                                   -> SB
#   Spontaneous Abortion -> SA; Induced Abortion and unspecified Abortion -> AB
#   Ectopic -> ECT
#   Molar -> AB (the HIP concept list has no molar codes, so the algorithm
#     cannot produce a molar category; pooled with AB and flagged)
#   Gravida / Para / *Unspecified -> PREG (recorded pregnancy, outcome not
#     resolved; matches the PMAP adapter's treatment of no-outcome rows)
edw_episodes <- edw_raw %>%
  mutate(
    pregnancy_outcome = as.character(pregnancy_outcome),
    has_delivery = suppressWarnings(as.numeric(has_delivery)),
    had_fetal_demise = suppressWarnings(as.numeric(had_fetal_demise)),
    n_births = coalesce(suppressWarnings(as.numeric(n_births)), 0),
    n_born_alive = coalesce(suppressWarnings(as.numeric(n_born_alive)), 0),
    n_fetal_demise = coalesce(suppressWarnings(as.numeric(n_fetal_demise)), 0),
    gt_start = coalesce(to_date_any(pregnancy_estimated_start_date),
                        to_date_any(episode_start_date)),
    gt_end = case_when(
      !is.na(has_delivery) & has_delivery == 1 & !is.na(to_date_any(last_delivery_date)) ~
        to_date_any(last_delivery_date),
      TRUE ~ coalesce(to_date_any(pregnancy_estimated_end_date),
                      to_date_any(episode_end_date))
    ),
    gt_ga_days = coalesce(suppressWarnings(as.numeric(last_delivery_gestational_age)),
                          suppressWarnings(as.numeric(ga_days_birth))),
    is_delivery_outcome = pregnancy_outcome %in% c("Term", "Preterm"),
    all_fetal_demise = (n_births > 0 & n_fetal_demise == n_births & n_born_alive == 0) |
      (n_births == 0 & coalesce(had_fetal_demise, 0) == 1),
    gt_outcome_category = case_when(
      is_delivery_outcome & all_fetal_demise ~ "SB",
      is_delivery_outcome ~ "LB",
      pregnancy_outcome == "Spontaneous Abortion" ~ "SA",
      pregnancy_outcome %in% c("Induced Abortion", "Abortion") ~ "AB",
      pregnancy_outcome == "Ectopic" ~ "ECT",
      pregnancy_outcome == "Molar" ~ "AB",
      TRUE ~ "PREG"
    ),
    gt_molar = pregnancy_outcome == "Molar",
    gt_unresolved = pregnancy_outcome %in% c("Gravida", "Para", "*Unspecified"),
    # Date quality: which fields were present, and whether the pair is usable.
    # An end before the start makes an interval that can never overlap an
    # algorithm episode, so keep only the end (a single-day window) and flag it.
    gt_start_source = case_when(
      !is.na(to_date_any(pregnancy_estimated_start_date)) ~ "estimated_start",
      !is.na(to_date_any(episode_start_date)) ~ "episode_start",
      TRUE ~ "none"
    ),
    gt_end_source = case_when(
      !is.na(has_delivery) & has_delivery == 1 & !is.na(to_date_any(last_delivery_date)) ~ "last_delivery",
      !is.na(to_date_any(pregnancy_estimated_end_date)) ~ "estimated_end",
      !is.na(to_date_any(episode_end_date)) ~ "episode_end",
      TRUE ~ "none"
    ),
    gt_dates_inconsistent = !is.na(gt_start) & !is.na(gt_end) & gt_end < gt_start,
    gt_start = if_else(gt_dates_inconsistent, as.Date(NA), gt_start)
  ) %>%
  filter(!is.na(person_id), !is.na(gt_start) | !is.na(gt_end)) %>%
  filter(is.na(gt_start) | gt_start < as.Date(params$max_start_date))

cat("\nEDW date quality by outcome (start source x end source, inconsistent = end before start):\n")
edw_episodes %>%
  count(pregnancy_outcome, gt_start_source, gt_end_source, gt_dates_inconsistent) %>%
  arrange(pregnancy_outcome, desc(n)) %>%
  tibble::as_tibble() %>%
  print(n = 60)
cat("\nEpisode duration (days) by outcome, from the dates as used:\n")
edw_episodes %>%
  mutate(dur = as.numeric(gt_end - gt_start)) %>%
  group_by(pregnancy_outcome) %>%
  summarise(n = n(), start_missing = sum(is.na(gt_start)), inconsistent = sum(gt_dates_inconsistent),
            median_dur = median(dur, na.rm = TRUE), pct_zero = round(100 * mean(dur == 0, na.rm = TRUE), 1),
            .groups = "drop") %>%
  arrange(desc(n)) %>%
  tibble::as_tibble() %>%
  print(n = 20)

gt_episodes <- edw_episodes %>%
  group_by(person_id) %>%
  arrange(coalesce(gt_end, gt_start), .by_group = TRUE) %>%
  mutate(gt_episode_num = row_number()) %>%
  ungroup() %>%
  select(person_id, cohort_id, pregnancy_key, gt_episode_num, gt_start, gt_end,
         gt_outcome_category, gt_ga_days, pregnancy_outcome, gt_molar, gt_unresolved,
         gt_start_source, gt_end_source, gt_dates_inconsistent)

cat("Derived outcomes for", nrow(gt_episodes), "ground truth episodes\n")
cat("Unique persons in ground truth:", n_distinct(gt_episodes$person_id), "\n")
cat("Unresolved (Gravida/Para/Unspecified) episodes mapped to PREG:",
    sum(gt_episodes$gt_unresolved), "\n")
cat("Molar pregnancies pooled with AB:", sum(gt_episodes$gt_molar), "\n")

algo_persons_set <- unique(algo_predictions$person_id)
gt_persons_set <- unique(gt_episodes$person_id)
cat("Person overlap (algo ∩ GT):", length(intersect(algo_persons_set, gt_persons_set)), "\n")

cat("\nGround truth outcome distribution (EDW PregnancyOutcome -> category):\n")
gt_episodes %>%
  count(pregnancy_outcome, gt_outcome_category) %>%
  arrange(desc(n)) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
```

```{r episode-level-matching}
# =========================================================
# Episode-level matching
# Primary analysis:
#   - outcome-agnostic matching
#   - exclude likely historical GT episodes
# Sensitivity analyses:
#   - include all GT episodes
#   - end-date windows (±14 / ±30 / ±60 days)
# =========================================================

prepare_validation_inputs <- function(algo_episodes,
                                      gt_episodes,
                                      historical_buffer_days = 30) {

  algo <- algo_episodes %>%
    dplyr::select(
      person_id,
      episode_number,
      episode_start_date,
      episode_end_date,
      outcome_category,
      gestational_age_days
    ) %>%
    dplyr::rename(
      algo_episode_num = episode_number,
      algo_outcome = outcome_category,
      algo_ga_days = gestational_age_days
    ) %>%
    dplyr::mutate(
      algo_start = as.Date(episode_start_date),
      algo_end   = as.Date(episode_end_date),
      algo_start = dplyr::coalesce(algo_start, algo_end),
      algo_end   = dplyr::coalesce(algo_end, algo_start)
    ) %>%
    dplyr::filter(!is.na(person_id), !is.na(algo_start), !is.na(algo_end)) %>%
    dplyr::select(person_id, algo_episode_num, algo_start, algo_end, algo_outcome, algo_ga_days) %>%
    dplyr::group_by(person_id) %>%
    dplyr::arrange(algo_end, .by_group = TRUE) %>%
    dplyr::mutate(algo_order = dplyr::row_number()) %>%
    dplyr::ungroup()

  gt_all <- gt_episodes %>%
    dplyr::select(
      person_id,
      gt_episode_num,
      gt_start,
      gt_end,
      gt_outcome_category,
      gt_ga_days
    ) %>%
    dplyr::rename(
      gt_outcome = gt_outcome_category
    ) %>%
    dplyr::mutate(
      gt_start = as.Date(gt_start),
      gt_end   = as.Date(gt_end),
      gt_start = dplyr::coalesce(gt_start, gt_end),
      gt_end   = dplyr::coalesce(gt_end, gt_start)
    ) %>%
    dplyr::filter(!is.na(person_id), !is.na(gt_start), !is.na(gt_end)) %>%
    dplyr::group_by(person_id) %>%
    dplyr::arrange(gt_end, .by_group = TRUE) %>%
    dplyr::mutate(gt_order = dplyr::row_number()) %>%
    dplyr::ungroup()

  # Person-specific observable anchor from algorithm episodes
  # GT episodes ending well before the first algorithm-observable episode
  # are treated as likely historical and excluded from the primary analysis.
  person_anchor <- algo %>%
    dplyr::group_by(person_id) %>%
    dplyr::summarise(
      first_algo_start = min(algo_start, na.rm = TRUE),
      last_algo_end = max(algo_end, na.rm = TRUE),
      .groups = "drop"
    )

  gt_all <- gt_all %>%
    dplyr::left_join(person_anchor, by = "person_id") %>%
    dplyr::mutate(
      likely_historical = dplyr::if_else(
        !is.na(first_algo_start) & gt_end < (first_algo_start - historical_buffer_days),
        TRUE,
        FALSE,
        missing = FALSE
      ),
      eligible_primary = !likely_historical
    )

  gt_primary <- gt_all %>%
    dplyr::filter(eligible_primary)

  historical_summary <- tibble::tibble(
    total_gt_episodes = nrow(gt_all),
    likely_historical_gt_episodes = sum(gt_all$likely_historical, na.rm = TRUE),
    eligible_gt_episodes = nrow(gt_primary),
    pct_historical = ifelse(nrow(gt_all) > 0,
                            100 * sum(gt_all$likely_historical, na.rm = TRUE) / nrow(gt_all),
                            NA_real_)
  )

  list(
    algo = algo,
    gt_all = gt_all,
    gt_primary = gt_primary,
    historical_summary = historical_summary
  )
}

build_match_candidates <- function(algo, gt) {

  common_persons <- intersect(unique(algo$person_id), unique(gt$person_id))

  cat("Persons in algorithm:", length(unique(algo$person_id)), "\n")
  cat("Persons in ground truth:", length(unique(gt$person_id)), "\n")
  cat("Person overlap for matching:", length(common_persons), "\n")

  if (length(common_persons) == 0) {
    warning("No person_id overlap between algorithm and ground truth.")
    return(tibble::tibble())
  }

  algo_sub <- algo %>%
    dplyr::filter(person_id %in% common_persons) %>%
    dplyr::mutate(.join_key = 1L)

  gt_sub <- gt %>%
    dplyr::filter(person_id %in% common_persons) %>%
    dplyr::mutate(.join_key = 1L)

  # Intentional many-to-many join:
  # within each person, create all possible algorithm-GT episode pairs
  candidates <- algo_sub %>%
    dplyr::inner_join(
      gt_sub,
      by = c("person_id", ".join_key"),
      relationship = "many-to-many"
    ) %>%
    dplyr::select(-.join_key) %>%
    dplyr::mutate(
      overlap_start = pmax(algo_start, gt_start),
      overlap_end   = pmin(algo_end, gt_end),
      overlap_days  = pmax(0, as.numeric(overlap_end - overlap_start) + 1),
      overlap_any   = overlap_days > 0,
      start_diff_days = abs(as.numeric(algo_start - gt_start)),
      end_diff_days   = abs(as.numeric(algo_end - gt_end)),
      mid_algo = algo_start + floor(as.numeric(algo_end - algo_start) / 2),
      mid_gt   = gt_start + floor(as.numeric(gt_end - gt_start) / 2),
      midpoint_diff_days = abs(as.numeric(mid_algo - mid_gt)),
      algo_key = paste(person_id, algo_episode_num, sep = "::"),
      gt_key   = paste(person_id, gt_episode_num, sep = "::")
    )

  candidates
}

greedy_one_to_one_match <- function(candidates,
                                    algo,
                                    gt,
                                    rule = c("overlap", "window"),
                                    date_window_days = 30) {

  rule <- match.arg(rule)

  if (nrow(candidates) == 0) {
    return(list(
      matched = tibble::tibble(),
      unmatched_algo = algo,
      unmatched_gt = gt,
      stats = list(
        total_algo = nrow(algo),
        total_gt = nrow(gt),
        matched = 0L,
        match_rate_algo = 0,
        match_rate_gt = 0
      )
    ))
  }

  candidates_ranked <- candidates

  if (rule == "overlap") {
    candidates_ranked <- candidates_ranked %>%
      dplyr::filter(overlap_any) %>%
      dplyr::arrange(
        person_id,
        dplyr::desc(overlap_days),
        end_diff_days,
        start_diff_days,
        midpoint_diff_days,
        algo_order,
        gt_order
      )
  }

  if (rule == "window") {
    candidates_ranked <- candidates_ranked %>%
      dplyr::filter(end_diff_days <= date_window_days) %>%
      dplyr::arrange(
        person_id,
        end_diff_days,
        start_diff_days,
        midpoint_diff_days,
        dplyr::desc(overlap_days),
        algo_order,
        gt_order
      )
  }

  if (nrow(candidates_ranked) == 0) {
    return(list(
      matched = tibble::tibble(),
      unmatched_algo = algo,
      unmatched_gt = gt,
      stats = list(
        total_algo = nrow(algo),
        total_gt = nrow(gt),
        matched = 0L,
        match_rate_algo = 0,
        match_rate_gt = 0
      )
    ))
  }

  used_algo <- character(0)
  used_gt <- character(0)
  keep <- logical(nrow(candidates_ranked))

  for (i in seq_len(nrow(candidates_ranked))) {
    this_algo <- candidates_ranked$algo_key[i]
    this_gt   <- candidates_ranked$gt_key[i]

    if (!(this_algo %in% used_algo) && !(this_gt %in% used_gt)) {
      keep[i] <- TRUE
      used_algo <- c(used_algo, this_algo)
      used_gt   <- c(used_gt, this_gt)
    }
  }

  matched <- candidates_ranked[keep, , drop = FALSE]

  unmatched_algo <- algo %>%
    dplyr::filter(!(paste(person_id, algo_episode_num, sep = "::") %in% matched$algo_key))

  unmatched_gt <- gt %>%
    dplyr::filter(!(paste(person_id, gt_episode_num, sep = "::") %in% matched$gt_key))

  stats <- list(
    total_algo = nrow(algo),
    total_gt = nrow(gt),
    matched = nrow(matched),
    match_rate_algo = ifelse(nrow(algo) > 0, nrow(matched) / nrow(algo), 0),
    match_rate_gt   = ifelse(nrow(gt) > 0, nrow(matched) / nrow(gt), 0)
  )

  list(
    matched = matched,
    unmatched_algo = unmatched_algo,
    unmatched_gt = unmatched_gt,
    stats = stats
  )
}

run_matching_analyses <- function(algo_episodes,
                                  gt_episodes,
                                  historical_buffer_days = 30) {

  prepared <- prepare_validation_inputs(
    algo_episodes = algo_episodes,
    gt_episodes = gt_episodes,
    historical_buffer_days = historical_buffer_days
  )

  algo <- prepared$algo
  gt_all <- prepared$gt_all
  gt_primary <- prepared$gt_primary

  # Primary analysis: eligible GT only
  candidates_primary <- build_match_candidates(algo, gt_primary)

  primary_overlap <- greedy_one_to_one_match(
    candidates = candidates_primary,
    algo = algo,
    gt = gt_primary,
    rule = "overlap"
  )

  end_window_14 <- greedy_one_to_one_match(
    candidates = candidates_primary,
    algo = algo,
    gt = gt_primary,
    rule = "window",
    date_window_days = 14
  )

  end_window_30 <- greedy_one_to_one_match(
    candidates = candidates_primary,
    algo = algo,
    gt = gt_primary,
    rule = "window",
    date_window_days = 30
  )

  end_window_60 <- greedy_one_to_one_match(
    candidates = candidates_primary,
    algo = algo,
    gt = gt_primary,
    rule = "window",
    date_window_days = 60
  )

  # Sensitivity: include all GT episodes
  candidates_all_gt <- build_match_candidates(algo, gt_all)

  all_gt_overlap <- greedy_one_to_one_match(
    candidates = candidates_all_gt,
    algo = algo,
    gt = gt_all,
    rule = "overlap"
  )

  summary_table <- tibble::tibble(
    matching_rule = c(
      "Primary: any episode overlap (eligible GT only)",
      "Sensitivity: any episode overlap (all GT episodes)",
      "Sensitivity: end date within ±14 days (eligible GT only)",
      "Sensitivity: end date within ±30 days (eligible GT only)",
      "Sensitivity: end date within ±60 days (eligible GT only)"
    ),
    matched_episodes = c(
      primary_overlap$stats$matched,
      all_gt_overlap$stats$matched,
      end_window_14$stats$matched,
      end_window_30$stats$matched,
      end_window_60$stats$matched
    ),
    algorithm_match_rate = c(
      primary_overlap$stats$match_rate_algo,
      all_gt_overlap$stats$match_rate_algo,
      end_window_14$stats$match_rate_algo,
      end_window_30$stats$match_rate_algo,
      end_window_60$stats$match_rate_algo
    ),
    ground_truth_match_rate = c(
      primary_overlap$stats$match_rate_gt,
      all_gt_overlap$stats$match_rate_gt,
      end_window_14$stats$match_rate_gt,
      end_window_30$stats$match_rate_gt,
      end_window_60$stats$match_rate_gt
    )
  )

  list(
    algo = algo,
    gt_all = gt_all,
    gt_primary = gt_primary,
    historical_summary = prepared$historical_summary,
    candidates_primary = candidates_primary,
    candidates_all_gt = candidates_all_gt,
    primary_overlap = primary_overlap,
    all_gt_overlap = all_gt_overlap,
    end_window_14 = end_window_14,
    end_window_30 = end_window_30,
    end_window_60 = end_window_60,
    summary_table = summary_table
  )
}

match_results <- run_matching_analyses(
  algo_episodes = algo_predictions,
  gt_episodes   = gt_episodes,
  historical_buffer_days = 30
)

matched_episodes <- match_results$primary_overlap$matched

cat("Historical GT Filtering Summary:\n")
cat("- Total GT episodes:", match_results$historical_summary$total_gt_episodes, "\n")
cat("- Likely historical GT episodes removed from primary analysis:",
    match_results$historical_summary$likely_historical_gt_episodes, "\n")
cat("- Eligible GT episodes for primary analysis:",
    match_results$historical_summary$eligible_gt_episodes, "\n")
cat("- Percent likely historical:",
    round(match_results$historical_summary$pct_historical, 1), "%\n\n")

cat("Episode Matching Results (primary overlap-based analysis):\n")
cat("- Algorithm episodes:", match_results$primary_overlap$stats$total_algo, "\n")
cat("- Eligible GT episodes:", match_results$primary_overlap$stats$total_gt, "\n")
cat("- Matched episodes:", match_results$primary_overlap$stats$matched, "\n")
cat("- Algorithm match rate:",
    round(100 * match_results$primary_overlap$stats$match_rate_algo, 1), "%\n")
cat("- Eligible GT match rate:",
    round(100 * match_results$primary_overlap$stats$match_rate_gt, 1), "%\n\n")

cat("Sensitivity Analysis (all GT episodes):\n")
cat("- All GT episodes:", match_results$all_gt_overlap$stats$total_gt, "\n")
cat("- Matched episodes:", match_results$all_gt_overlap$stats$matched, "\n")
cat("- Algorithm match rate:",
    round(100 * match_results$all_gt_overlap$stats$match_rate_algo, 1), "%\n")
cat("- All-GT match rate:",
    round(100 * match_results$all_gt_overlap$stats$match_rate_gt, 1), "%\n")
```

## Matching summary table
```{r historical-gt-summary-table}
match_results$summary_table %>%
  dplyr::mutate(
    algorithm_match_rate = round(100 * algorithm_match_rate, 1),
    ground_truth_match_rate = round(100 * ground_truth_match_rate, 1)
  ) %>%
  knitr::kable(
    caption = "Episode Matching Summary Across Primary and Sensitivity Analyses",
    col.names = c(
      "Matching Rule",
      "Matched Episodes",
      "Algorithm Match Rate (%)",
      "Ground Truth Match Rate (%)"
    )
  ) %>%
  kableExtra::kable_styling(
    bootstrap_options = c("striped", "hover"),
    full_width = FALSE
  )
```



# Outcome classification analysis
```{r outcome-classification-exploratory}
# =========================================================
# Exploratory outcome classification analysis
# Labels are not yet finalized, so this is reported as
# exploratory only and is not part of the primary endpoint.
# =========================================================

create_confusion_matrix <- function(matched_episodes) {

  outcome_order <- c("LB", "SB", "ECT", "AB", "SA")

  conf_matrix <- matched_episodes %>%
    dplyr::filter(
      algo_outcome %in% outcome_order,
      gt_outcome %in% outcome_order
    ) %>%
    dplyr::mutate(
      algo_outcome = factor(algo_outcome, levels = outcome_order),
      gt_outcome   = factor(gt_outcome, levels = outcome_order)
    ) %>%
    dplyr::count(gt_outcome, algo_outcome, .drop = FALSE) %>%
    tidyr::pivot_wider(
      names_from = algo_outcome,
      values_from = n,
      values_fill = 0
    ) %>%
    dplyr::arrange(gt_outcome)

  conf_matrix
}

calculate_classification_metrics <- function(matched_episodes) {

  outcome_order <- c("LB", "SB", "ECT", "AB", "SA")

  dat <- matched_episodes %>%
    dplyr::filter(
      algo_outcome %in% outcome_order,
      gt_outcome %in% outcome_order
    )

  if (nrow(dat) == 0) {
    return(list(
      overall_accuracy = NA_real_,
      total_correct = 0L,
      total_matched = 0L,
      by_category = tibble::tibble()
    ))
  }

  metrics_by_category <- purrr::map_dfr(outcome_order, function(cat) {

    tp <- sum(dat$algo_outcome == cat & dat$gt_outcome == cat, na.rm = TRUE)
    fp <- sum(dat$algo_outcome == cat & dat$gt_outcome != cat, na.rm = TRUE)
    fn <- sum(dat$algo_outcome != cat & dat$gt_outcome == cat, na.rm = TRUE)

    precision <- ifelse(tp + fp > 0, tp / (tp + fp), NA_real_)
    recall    <- ifelse(tp + fn > 0, tp / (tp + fn), NA_real_)
    f1_score  <- ifelse(
      !is.na(precision) && !is.na(recall) && (precision + recall) > 0,
      2 * precision * recall / (precision + recall),
      NA_real_
    )

    tibble::tibble(
      category = cat,
      support = sum(dat$gt_outcome == cat, na.rm = TRUE),
      predicted = sum(dat$algo_outcome == cat, na.rm = TRUE),
      tp = tp,
      fp = fp,
      fn = fn,
      precision = precision,
      recall = recall,
      f1_score = f1_score
    )
  })

  total_correct <- sum(dat$algo_outcome == dat$gt_outcome, na.rm = TRUE)
  total_matched <- nrow(dat)
  overall_accuracy <- ifelse(total_matched > 0, total_correct / total_matched, NA_real_)

  list(
    overall_accuracy = overall_accuracy,
    total_correct = total_correct,
    total_matched = total_matched,
    by_category = metrics_by_category
  )
}

conf_matrix <- create_confusion_matrix(matched_episodes)
class_metrics <- calculate_classification_metrics(matched_episodes)

cat("Exploratory Outcome Classification Results:\n")
cat("- This analysis is exploratory only and excludes unresolved categories.\n")
cat("- Overall accuracy:",
    round(100 * class_metrics$overall_accuracy, 1), "%\n")
cat("(", class_metrics$total_correct, "/", class_metrics$total_matched, "episodes with finalized labels)\n\n")
```

## Confusion Matrix
```{r confusion-matrix}

if (nrow(conf_matrix) > 0) {
  conf_matrix %>%
    knitr::kable(caption = "Exploratory Confusion Matrix: Ground Truth vs Algorithm Outcome") %>%
    kableExtra::kable_styling(
      bootstrap_options = c("striped", "hover"),
      full_width = FALSE
    )
}
```


## Classification metrics by category
```{r classifcation-metrics}
if (nrow(class_metrics$by_category) > 0) {
  class_metrics$by_category %>%
    dplyr::mutate(
      precision = round(precision, 3),
      recall = round(recall, 3),
      f1_score = round(f1_score, 3)
    ) %>%
    knitr::kable(
      caption = "Exploratory Classification Metrics by Finalized Outcome Category"
    ) %>%
    kableExtra::kable_styling(
      bootstrap_options = c("striped", "hover"),
      full_width = FALSE
    )
}
```

# Gestational age comparison

```{r gestational-age}
# =========================================================
# Gestational age comparison
# Primary analysis: overlap-matched episodes using eligible GT only
# =========================================================

calculate_ga_metrics <- function(matched_episodes, min_ga = 0, max_ga = 320) {

  with_ga <- matched_episodes %>%
    dplyr::filter(
      !is.na(algo_ga_days),
      !is.na(gt_ga_days),
      algo_ga_days >= min_ga,
      algo_ga_days <= max_ga,
      gt_ga_days >= min_ga,
      gt_ga_days <= max_ga
    ) %>%
    dplyr::mutate(
      ga_diff = algo_ga_days - gt_ga_days,
      abs_ga_diff = abs(ga_diff)
    )

  if (nrow(with_ga) == 0) {
    return(list(
      summary = tibble::tibble(),
      by_outcome = tibble::tibble(),
      data = with_ga
    ))
  }

  summary <- with_ga %>%
    dplyr::summarise(
      n = dplyr::n(),
      mean_diff_days = mean(ga_diff, na.rm = TRUE),
      median_diff_days = stats::median(ga_diff, na.rm = TRUE),
      mean_abs_diff_days = mean(abs_ga_diff, na.rm = TRUE),
      median_abs_diff_days = stats::median(abs_ga_diff, na.rm = TRUE),
      within_7_days = mean(abs_ga_diff <= 7, na.rm = TRUE),
      within_14_days = mean(abs_ga_diff <= 14, na.rm = TRUE),
      within_21_days = mean(abs_ga_diff <= 21, na.rm = TRUE)
    )

  by_outcome <- with_ga %>%
    dplyr::group_by(gt_outcome) %>%
    dplyr::summarise(
      n = dplyr::n(),
      mean_diff_days = mean(ga_diff, na.rm = TRUE),
      median_diff_days = stats::median(ga_diff, na.rm = TRUE),
      mean_abs_diff_days = mean(abs_ga_diff, na.rm = TRUE),
      within_7_days = mean(abs_ga_diff <= 7, na.rm = TRUE),
      within_14_days = mean(abs_ga_diff <= 14, na.rm = TRUE),
      .groups = "drop"
    ) %>%
    dplyr::arrange(dplyr::desc(n))

  list(
    summary = summary,
    by_outcome = by_outcome,
    data = with_ga
  )
}

ga_metrics <- calculate_ga_metrics(matched_episodes)

if (nrow(ga_metrics$summary) > 0) {
  cat("Gestational Age Results (primary overlap-matched episodes):\n")
  cat("- Mean GA difference:", round(ga_metrics$summary$mean_diff_days, 1), "days\n")
  cat("- Median GA difference:", round(ga_metrics$summary$median_diff_days, 1), "days\n")
  cat("- Mean absolute GA difference:", round(ga_metrics$summary$mean_abs_diff_days, 1), "days\n")
  cat("- Within ±7 days:", round(100 * ga_metrics$summary$within_7_days, 1), "%\n")
  cat("- Within ±14 days:", round(100 * ga_metrics$summary$within_14_days, 1), "%\n")
  cat("- Within ±21 days:", round(100 * ga_metrics$summary$within_21_days, 1), "%\n")
}
```

## GA summary table
```{r GA-summary}
if (nrow(ga_metrics$summary) > 0) {
  ga_metrics$summary %>%
    dplyr::mutate(dplyr::across(where(is.numeric), ~ round(., 2))) %>%
    tidyr::pivot_longer(
      everything(),
      names_to = "Metric",
      values_to = "Value"
    ) %>%
    knitr::kable(caption = "Gestational Age Comparison Summary") %>%
    kableExtra::kable_styling(
      bootstrap_options = c("striped", "hover"),
      full_width = FALSE
    )
}
```


## GA by outcome category
```{r GA-outcome-category}
if (nrow(ga_metrics$by_outcome) > 0) {
  ga_metrics$by_outcome %>%
    dplyr::mutate(
      dplyr::across(
        where(is.numeric) & !dplyr::matches("^n$"),
        ~ round(., 2)
      )
    ) %>%
    knitr::kable(caption = "GA Agreement by Ground Truth Outcome Category") %>%
    kableExtra::kable_styling(
      bootstrap_options = c("striped", "hover"),
      full_width = FALSE
    )
}
```


## GA agreement plots
```{r GA-agreement}
if (nrow(ga_metrics$data) > 0) {
  ggplot2::ggplot(
    ga_metrics$data,
    ggplot2::aes(x = gt_ga_days, y = algo_ga_days)
  ) +
    ggplot2::geom_point(alpha = 0.3, size = 1.5) +
    ggplot2::geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "red") +
    ggplot2::labs(
      title = "Predicted vs Ground Truth Gestational Age",
      x = "Ground Truth GA (days)",
      y = "Algorithm GA (days)"
    ) +
    ggplot2::theme_minimal()
}
```


## GA metrics plot
```{r GA-bland-altman-plot}
if (nrow(ga_metrics$data) > 0) {

  ba_data <- ga_metrics$data %>%
    dplyr::mutate(
      mean_ga = (algo_ga_days + gt_ga_days) / 2,
      diff_ga = algo_ga_days - gt_ga_days
    ) %>%
    dplyr::filter(
      !is.na(mean_ga),
      !is.na(diff_ga)
    )

  if (nrow(ba_data) >= 2) {

    mean_diff <- mean(ba_data$diff_ga, na.rm = TRUE)
    sd_diff <- stats::sd(ba_data$diff_ga, na.rm = TRUE)

    upper_loa <- mean_diff + 1.96 * sd_diff
    lower_loa <- mean_diff - 1.96 * sd_diff

    ggplot2::ggplot(
      ba_data,
      ggplot2::aes(x = mean_ga, y = diff_ga)
    ) +
      ggplot2::geom_point(alpha = 0.3, size = 1.5) +
      ggplot2::geom_hline(
        yintercept = mean_diff,
        linetype = "solid",
        color = "red"
      ) +
      ggplot2::geom_hline(
        yintercept = upper_loa,
        linetype = "dashed",
        color = "gray40"
      ) +
      ggplot2::geom_hline(
        yintercept = lower_loa,
        linetype = "dashed",
        color = "gray40"
      ) +
      ggplot2::annotate(
        "text",
        x = Inf,
        y = mean_diff,
        label = paste0("Mean bias = ", round(mean_diff, 1)),
        hjust = 1.1,
        vjust = -0.5,
        size = 3.5,
        color = "red"
      ) +
      ggplot2::annotate(
        "text",
        x = Inf,
        y = upper_loa,
        label = paste0("+1.96 SD = ", round(upper_loa, 1)),
        hjust = 1.1,
        vjust = -0.5,
        size = 3.5,
        color = "gray40"
      ) +
      ggplot2::annotate(
        "text",
        x = Inf,
        y = lower_loa,
        label = paste0("-1.96 SD = ", round(lower_loa, 1)),
        hjust = 1.1,
        vjust = 1.5,
        size = 3.5,
        color = "gray40"
      ) +
      ggplot2::labs(
        title = "Bland-Altman Plot for Gestational Age",
        x = "Mean GA (days)",
        y = "Algorithm - Ground Truth GA (days)"
      ) +
      ggplot2::theme_minimal()

  } else {
    cat("Not enough non-missing observations to generate a Bland-Altman plot.\n")
  }
}
```
```{r plot-tests}
ggplot2::ggplot(
  ga_metrics$data,
  ggplot2::aes(x = gt_ga_days, y = algo_ga_days)
) +
  ggplot2::geom_point(alpha = 0.3, size = 1.5) +
  ggplot2::geom_abline(
    slope = 1,
    intercept = 0,
    linetype = "dashed",
    color = "red"
  ) +
  ggplot2::labs(
    title = "Algorithm vs Ground Truth Gestational Age",
    x = "Ground Truth GA (days)",
    y = "Algorithm GA (days)"
  ) +
  ggplot2::theme_minimal()

ga_metrics$data %>%
  dplyr::mutate(diff_ga = algo_ga_days - gt_ga_days) %>%
  ggplot2::ggplot(
    ggplot2::aes(x = diff_ga)
  ) +
  ggplot2::geom_density(fill = "steelblue", alpha = 0.5) +
  ggplot2::geom_vline(xintercept = 0, linetype = "dashed") +
  ggplot2::labs(
    title = "Distribution of Gestational Age Differences",
    x = "Algorithm − Ground Truth GA (days)",
    y = "Density"
  ) +
  ggplot2::theme_minimal()

ga_metrics$data %>%
  dplyr::mutate(
    abs_diff = abs(algo_ga_days - gt_ga_days)
  ) %>%
  ggplot2::ggplot(
    ggplot2::aes(x = abs_diff)
  ) +
  ggplot2::stat_ecdf() +
  ggplot2::labs(
    title = "Cumulative Accuracy of Gestational Age Estimates",
    x = "Absolute Difference (days)",
    y = "Proportion of Episodes"
  ) +
  ggplot2::theme_minimal()

ggplot2::ggplot(
  ga_metrics$data,
  ggplot2::aes(x = gt_ga_days, y = algo_ga_days)
) +
  ggplot2::geom_bin2d() +
  ggplot2::scale_fill_viridis_c() +
  ggplot2::labs(
    title = "Density of Gestational Age Agreement",
    x = "Ground Truth GA (days)",
    y = "Algorithm GA (days)"
  ) +
  ggplot2::theme_minimal()
```

```{r plots-test2}
ggplot2::ggplot(
  ga_metrics$data,
  ggplot2::aes(x = gt_ga_days, y = algo_ga_days)
) +
  ggplot2::geom_hex() +
  ggplot2::scale_fill_viridis_c() +
  ggplot2::geom_abline(slope = 1, intercept = 0, linetype = "dashed") +
  ggplot2::labs(
    title = "Density of Gestational Age Agreement",
    x = "Ground Truth GA (days)",
    y = "Algorithm GA (days)"
  ) +
  ggplot2::theme_minimal()

ga_metrics$data %>%
  dplyr::mutate(
    diff_ga = algo_ga_days - gt_ga_days
  ) %>%
  ggplot2::ggplot(
    ggplot2::aes(y = diff_ga)
  ) +
  ggplot2::geom_boxplot(fill = "steelblue") +
  ggplot2::geom_hline(yintercept = 0, linetype = "dashed") +
  ggplot2::labs(
    title = "Distribution of Gestational Age Error",
    y = "Algorithm − Ground Truth (days)"
  ) +
  ggplot2::theme_minimal()


ga_metrics$data %>%
  dplyr::mutate(
    diff_ga = algo_ga_days - gt_ga_days
  ) %>%
  ggplot2::ggplot(
    ggplot2::aes(x = "", y = diff_ga)
  ) +
  ggplot2::geom_violin(fill = "steelblue", alpha = 0.6) +
  ggplot2::geom_hline(yintercept = 0, linetype = "dashed") +
  ggplot2::labs(
    title = "Distribution of Gestational Age Differences",
    x = "",
    y = "Algorithm − Ground Truth (days)"
  ) +
  ggplot2::theme_minimal()


ga_metrics$data %>%
  dplyr::mutate(
    abs_diff = abs(algo_ga_days - gt_ga_days),
    category = dplyr::case_when(
      abs_diff <= 7 ~ "≤7 days",
      abs_diff <= 14 ~ "8–14 days",
      abs_diff <= 30 ~ "15–30 days",
      TRUE ~ ">30 days"
    )
  ) %>%
  ggplot2::ggplot(
    ggplot2::aes(x = category)
  ) +
  ggplot2::geom_bar(fill = "steelblue") +
  ggplot2::labs(
    title = "Accuracy Categories of Gestational Age Estimation",
    x = "Absolute Difference",
    y = "Number of Episodes"
  ) +
  ggplot2::theme_minimal()


```

```{r bland-altman-new}
if (nrow(ga_metrics$data) > 0) {

  ba_data <- ga_metrics$data %>%
    dplyr::mutate(
      mean_ga = (algo_ga_days + gt_ga_days) / 2,
      diff_ga = algo_ga_days - gt_ga_days
    ) %>%
    dplyr::filter(
      !is.na(mean_ga),
      !is.na(diff_ga)
    )

  if (nrow(ba_data) >= 2) {

    mean_diff <- mean(ba_data$diff_ga, na.rm = TRUE)
    sd_diff <- stats::sd(ba_data$diff_ga, na.rm = TRUE)

    upper_loa <- mean_diff + 1.96 * sd_diff
    lower_loa <- mean_diff - 1.96 * sd_diff

    x_max <- max(ba_data$mean_ga, na.rm = TRUE)
    x_min <- min(ba_data$mean_ga, na.rm = TRUE)
    x_pad <- 0.03 * (x_max - x_min)

    ggplot2::ggplot(
      ba_data,
      ggplot2::aes(x = mean_ga, y = diff_ga)
    ) +
      ggplot2::annotate(
        "rect",
        xmin = -Inf, xmax = Inf,
        ymin = lower_loa, ymax = upper_loa,
        fill = "grey85",
        alpha = 0.35
      ) +
      ggplot2::geom_hline(
        yintercept = 0,
        linetype = "dotted",
        linewidth = 0.5,
        color = "grey50"
      ) +
      ggplot2::geom_hline(
        yintercept = mean_diff,
        linewidth = 0.9,
        color = "#C44E52"
      ) +
      ggplot2::geom_hline(
        yintercept = upper_loa,
        linetype = "dashed",
        linewidth = 0.7,
        color = "grey35"
      ) +
      ggplot2::geom_hline(
        yintercept = lower_loa,
        linetype = "dashed",
        linewidth = 0.7,
        color = "grey35"
      ) +
      ggplot2::geom_point(
        alpha = 0.28,
        size = 2,
        color = "#3B76AF"
      ) +
      ggplot2::geom_smooth(
        method = "loess",
        se = FALSE,
        linewidth = 0.9,
        color = "#2F4B5C"
      ) +
      ggplot2::annotate(
        "label",
        x = x_max + x_pad,
        y = mean_diff,
        label = paste0("Mean bias = ", round(mean_diff, 1)),
        hjust = 0,
        vjust = -0.2,
        size = 3.6,
        label.size = 0.2,
        fill = "white",
        color = "#C44E52"
      ) +
      ggplot2::annotate(
        "label",
        x = x_max + x_pad,
        y = upper_loa,
        label = paste0("Upper LOA = ", round(upper_loa, 1)),
        hjust = 0,
        vjust = -0.2,
        size = 3.4,
        label.size = 0.2,
        fill = "white",
        color = "grey25"
      ) +
      ggplot2::annotate(
        "label",
        x = x_max + x_pad,
        y = lower_loa,
        label = paste0("Lower LOA = ", round(lower_loa, 1)),
        hjust = 0,
        vjust = 1.2,
        size = 3.4,
        label.size = 0.2,
        fill = "white",
        color = "grey25"
      ) +
      ggplot2::scale_x_continuous(
        expand = ggplot2::expansion(mult = c(0.03, 0.22))
      ) +
      ggplot2::labs(
        title = "Bland-Altman Plot for Gestational Age",
        subtitle = "Difference between algorithm-derived and ground-truth gestational age",
        x = "Mean gestational age (days)",
        y = "Algorithm − Ground truth (days)",
        caption = paste0(
          "Shaded area indicates 95% limits of agreement. ",
          "N = ", nrow(ba_data)
        )
      ) +
      ggplot2::coord_cartesian(clip = "off") +
      ggplot2::theme_minimal(base_size = 12) +
      ggplot2::theme(
        plot.title = ggplot2::element_text(face = "bold", size = 14),
        plot.subtitle = ggplot2::element_text(size = 11, color = "grey30"),
        axis.title = ggplot2::element_text(face = "bold"),
        panel.grid.minor = ggplot2::element_blank(),
        panel.grid.major.x = ggplot2::element_line(color = "grey90"),
        panel.grid.major.y = ggplot2::element_line(color = "grey90"),
        plot.caption = ggplot2::element_text(color = "grey35", hjust = 0),
        plot.margin = ggplot2::margin(t = 10, r = 70, b = 10, l = 10)
      )

  } else {
    cat("Not enough non-missing observations to generate a Bland-Altman plot.\n")
  }
}
```

# Bland-Altman Check

## 1. Extract the suspicious region (manual inspection)
```{r ba-check1}
suspicious_cases <- ba_data %>%
  dplyr::filter(
    mean_ga < 180,
    diff_ga < -100
  )

nrow(suspicious_cases)

head(suspicious_cases)

suspicious_cases %>%
  dplyr::select(
    person_id,
    algo_episode_num,
    gt_episode_num,
    algo_ga_days,
    gt_ga_days,
    mean_ga,
    diff_ga
  ) %>%
  head(50)
```


## 2. whether the pattern is caused by GT truncation
```{r ba-check2}
summary(ba_data$gt_ga_days)
summary(ba_data$algo_ga_days)
hist(ba_data$gt_ga_days, breaks = 50)
hist(ba_data$algo_ga_days, breaks = 50)
```

## 3. N/A

## 4. Check if the pattern is driven by episode type
```{r ba-check4}
ba_data %>%
  dplyr::mutate(
    abs_diff = abs(diff_ga)
  ) %>%
  dplyr::group_by(gt_outcome) %>%
  dplyr::summarise(
    n = dplyr::n(),
    mean_diff = mean(diff_ga, na.rm = TRUE),
    large_error = mean(abs_diff > 100)
  )
```

## 5. Check whether historical pregnancies cause it
```{r ba-check5}

suspicious_cases <- ba_data %>%
  dplyr::filter(
    mean_ga < 180,
    diff_ga < -100
  )

nrow(suspicious_cases)

head(suspicious_cases, 20)

```

## 6. Visual diagnostic: highlight the suspicious region
```{r ba-check6}

ggplot2::ggplot(
  ba_data,
  ggplot2::aes(mean_ga, diff_ga)
) +
  ggplot2::geom_point(alpha = 0.2) +
  ggplot2::geom_point(
    data = suspicious_cases,
    color = "red",
    alpha = 0.6
  ) +
  ggplot2::labs(
    title = "Highlighted Suspicious Region"
  )
```


## 7. Check if GA calculation is inconsistent
```{r ba-check7}

cor(
  ba_data$algo_ga_days,
  ba_data$gt_ga_days,
  use = "complete.obs"
)
```

## 8. Plot GA difference vs GT GA
```{r ba-check8}
ggplot2::ggplot(
  ba_data,
  ggplot2::aes(
    x = gt_ga_days,
    y = diff_ga
  )
) +
  ggplot2::geom_point(alpha = 0.2) +
  ggplot2::geom_smooth()
```



# Personal-level aggregation
```{r person-level-aggregation}
# =========================================================
# Person-level aggregation
# Report both:
#   - primary comparison against eligible GT only
#   - sensitivity comparison against all GT episodes
# =========================================================

calculate_person_metrics <- function(algo_episodes, gt_episodes) {

  algo_counts <- algo_episodes %>%
    dplyr::group_by(person_id) %>%
    dplyr::summarise(
      algo_total = dplyr::n(),
      .groups = "drop"
    )

  gt_counts <- gt_episodes %>%
    dplyr::group_by(person_id) %>%
    dplyr::summarise(
      gt_total = dplyr::n(),
      .groups = "drop"
    )

  person_data <- dplyr::full_join(algo_counts, gt_counts, by = "person_id") %>%
    dplyr::mutate(
      algo_total = dplyr::coalesce(algo_total, 0L),
      gt_total   = dplyr::coalesce(gt_total, 0L),
      total_diff = algo_total - gt_total,
      abs_total_diff = abs(total_diff),
      exact_match = algo_total == gt_total,
      within_1 = abs_total_diff <= 1
    )

  summary <- person_data %>%
    dplyr::summarise(
      n_persons = dplyr::n(),
      exact_match_rate = mean(exact_match, na.rm = TRUE),
      within_1_rate = mean(within_1, na.rm = TRUE),
      mean_abs_diff = mean(abs_total_diff, na.rm = TRUE),
      median_abs_diff = stats::median(abs_total_diff, na.rm = TRUE)
    )

  list(
    summary = summary,
    person_data = person_data
  )
}

person_metrics_primary <- calculate_person_metrics(
  algo_episodes = algo_predictions,
  gt_episodes   = match_results$gt_primary
)

person_metrics_all_gt <- calculate_person_metrics(
  algo_episodes = algo_predictions,
  gt_episodes   = match_results$gt_all
)

cat("Person-Level Aggregation Results (primary eligible GT analysis):\n")
cat("- Exact count match:",
    round(100 * person_metrics_primary$summary$exact_match_rate, 1), "%\n")
cat("- Within ±1 episode:",
    round(100 * person_metrics_primary$summary$within_1_rate, 1), "%\n")
cat("- Mean absolute count difference:",
    round(person_metrics_primary$summary$mean_abs_diff, 2), "\n")
cat("- Median absolute count difference:",
    round(person_metrics_primary$summary$median_abs_diff, 2), "\n\n")

cat("Person-Level Aggregation Sensitivity (all GT episodes):\n")
cat("- Exact count match:",
    round(100 * person_metrics_all_gt$summary$exact_match_rate, 1), "%\n")
cat("- Within ±1 episode:",
    round(100 * person_metrics_all_gt$summary$within_1_rate, 1), "%\n")
cat("- Mean absolute count difference:",
    round(person_metrics_all_gt$summary$mean_abs_diff, 2), "\n")
cat("- Median absolute count difference:",
    round(person_metrics_all_gt$summary$median_abs_diff, 2), "\n")
```


## Person-level summary table
```{r person-level-summary}
person_summary_table <- tibble::tibble(
  analysis = c("Primary: eligible GT only", "Sensitivity: all GT episodes"),
  exact_match_rate = c(
    person_metrics_primary$summary$exact_match_rate,
    person_metrics_all_gt$summary$exact_match_rate
  ),
  within_1_rate = c(
    person_metrics_primary$summary$within_1_rate,
    person_metrics_all_gt$summary$within_1_rate
  ),
  mean_abs_diff = c(
    person_metrics_primary$summary$mean_abs_diff,
    person_metrics_all_gt$summary$mean_abs_diff
  ),
  median_abs_diff = c(
    person_metrics_primary$summary$median_abs_diff,
    person_metrics_all_gt$summary$median_abs_diff
  )
)

person_summary_table %>%
  dplyr::mutate(
    exact_match_rate = round(100 * exact_match_rate, 1),
    within_1_rate = round(100 * within_1_rate, 1),
    mean_abs_diff = round(mean_abs_diff, 2),
    median_abs_diff = round(median_abs_diff, 2)
  ) %>%
  knitr::kable(
    caption = "Person-Level Aggregation Summary",
    col.names = c(
      "Analysis",
      "Exact Match Rate (%)",
      "Within ±1 Episode (%)",
      "Mean Absolute Difference",
      "Median Absolute Difference"
    )
  ) %>%
  kableExtra::kable_styling(
    bootstrap_options = c("striped", "hover"),
    full_width = FALSE
  )
```


## Distribution of count differences
```{r dist-cnt-diff}
if (nrow(person_metrics_primary$person_data) > 0) {
  ggplot2::ggplot(
    person_metrics_primary$person_data,
    ggplot2::aes(x = total_diff)
  ) +
    ggplot2::geom_histogram(
      binwidth = 1,
      fill = "steelblue",
      color = "white"
    ) +
    ggplot2::labs(
      title = "Distribution of Person-Level Episode Count Differences (Eligible GT Only)",
      x = "Algorithm count - Ground truth count",
      y = "Number of persons"
    ) +
    ggplot2::theme_minimal()
}
```

# Summary and key findings
```{r summary-key-findings}
# =========================================================
# Summary and key findings
# Includes a podium-oriented subset
# =========================================================

summary_df <- tibble::tibble(
  Metric = c(
    "Algorithm episodes",
    "Ground truth episodes (all)",
    "Ground truth episodes (eligible for primary analysis)",
    "Likely historical GT episodes removed",
    "Primary matched episodes (overlap; eligible GT only)",
    "Algorithm match rate (primary eligible GT analysis)",
    "Eligible GT match rate (primary analysis)",
    "Algorithm match rate (all GT sensitivity analysis)",
    "All-GT match rate (sensitivity analysis)",
    "Mean GA difference (days)",
    "GA within ±7 days",
    "GA within ±14 days",
    "Person exact count match (eligible GT only)",
    "Person count within ±1 episode (eligible GT only)"
  ),
  Value = c(
    match_results$primary_overlap$stats$total_algo,
    match_results$all_gt_overlap$stats$total_gt,
    match_results$primary_overlap$stats$total_gt,
    match_results$historical_summary$likely_historical_gt_episodes,
    match_results$primary_overlap$stats$matched,
    round(100 * match_results$primary_overlap$stats$match_rate_algo, 1),
    round(100 * match_results$primary_overlap$stats$match_rate_gt, 1),
    round(100 * match_results$all_gt_overlap$stats$match_rate_algo, 1),
    round(100 * match_results$all_gt_overlap$stats$match_rate_gt, 1),
    ifelse(nrow(ga_metrics$summary) > 0, round(ga_metrics$summary$mean_diff_days, 1), NA_real_),
    ifelse(nrow(ga_metrics$summary) > 0, round(100 * ga_metrics$summary$within_7_days, 1), NA_real_),
    ifelse(nrow(ga_metrics$summary) > 0, round(100 * ga_metrics$summary$within_14_days, 1), NA_real_),
    round(100 * person_metrics_primary$summary$exact_match_rate, 1),
    round(100 * person_metrics_primary$summary$within_1_rate, 1)
  )
)

summary_df %>%
  knitr::kable(caption = "Overall Validation Summary") %>%
  kableExtra::kable_styling(
    bootstrap_options = c("striped", "hover"),
    full_width = FALSE
  )
```



## Podium abstract subset:
```{r podium-abstract}
# Podium-oriented subset:
# emphasize episode identification and gestational-age agreement
podium_df <- tibble::tibble(
  Metric = c(
    "Algorithm episodes",
    "Eligible GT episodes",
    "Primary matched episodes (overlap)",
    "Algorithm match rate (primary analysis)",
    "Eligible GT match rate (primary analysis)",
    "Mean GA difference (days)",
    "GA within ±7 days",
    "GA within ±14 days"
  ),
  Value = c(
    match_results$primary_overlap$stats$total_algo,
    match_results$primary_overlap$stats$total_gt,
    match_results$primary_overlap$stats$matched,
    round(100 * match_results$primary_overlap$stats$match_rate_algo, 1),
    round(100 * match_results$primary_overlap$stats$match_rate_gt, 1),
    ifelse(nrow(ga_metrics$summary) > 0, round(ga_metrics$summary$mean_diff_days, 1), NA_real_),
    ifelse(nrow(ga_metrics$summary) > 0, round(100 * ga_metrics$summary$within_7_days, 1), NA_real_),
    ifelse(nrow(ga_metrics$summary) > 0, round(100 * ga_metrics$summary$within_14_days, 1), NA_real_)
  )
)

podium_df %>%
  knitr::kable(caption = "Podium-Focused Result Subset") %>%
  kableExtra::kable_styling(
    bootstrap_options = c("striped", "hover"),
    full_width = FALSE
  )
```


## Key findings
```{r key-findings}
cat("Key Findings:\n")
cat(
  "1. After excluding likely historical ground-truth episodes from the primary analysis,",
  match_results$primary_overlap$stats$matched,
  "episodes were matched using outcome-agnostic overlap-based matching.\n"
)

cat(
  "2. In the primary analysis, the algorithm-side match rate was",
  round(100 * match_results$primary_overlap$stats$match_rate_algo, 1),
  "% and the eligible ground-truth-side match rate was",
  round(100 * match_results$primary_overlap$stats$match_rate_gt, 1),
  "%.\n"
)

cat(
  "3. Gestational age agreement among primary matched episodes was",
  ifelse(
    nrow(ga_metrics$summary) > 0,
    paste0(round(100 * ga_metrics$summary$within_14_days, 1), "% within ±14 days"),
    "not available"
  ),
  ".\n"
)

cat(
  "4. Exploratory outcome classification was not used as a primary endpoint because the outcome labels are not yet finalized.\n"
)

cat(
  "5. Person-level exact episode-count agreement was",
  round(100 * person_metrics_primary$summary$exact_match_rate, 1),
  "% in the primary eligible-GT analysis and",
  round(100 * person_metrics_all_gt$summary$exact_match_rate, 1),
  "% when all GT episodes were included.\n"
)

cat(
  "6. The difference between the primary and all-GT analyses quantifies the impact of likely historical registry episodes that may not be observable in the Hopkins healthcare system data used by the algorithm.\n"
)
```
```{r cleanup1}
#| include: false

# Disconnect from database
if (exists("con")) {
  db_disconnect(con)
  cat("Database connection closed\n")
}
```


# Pregnancy complication

## setup
```{r pregnancy-complications-setup}
# =========================================================
# Pregnancy complications detected by either approach
# New section:
#   - preeclampsia
#   - pregnancy-induced hypertension
#   - gestational diabetes
# =========================================================

complication_concepts <- tibble::tibble(
  complication = c(
    "Preeclampsia",
    "Pregnancy-induced hypertension",
    "Gestational diabetes"
  ),
  concept_id = c(
    439393,
    4167493,
    4024659
  ),
  short_name = c(
    "preeclampsia",
    "pih",
    "gdm"
  )
)
```

```{r db-connect2}
# Connect to database using project connection function
con <- open_report_connection()
if (is.null(con)) stop("Connection returned NULL.")
if (!db_is_valid(con)) stop("Database connection is not valid.")

cat("Database connection established\n")

```


```{r pregnancy-complications-descendants}
# =========================================================
# Step 1A: get all descendant concept IDs from OMOP
# =========================================================

ensure_connection <- function(conn) {
  ok <- tryCatch(db_is_valid(conn), error = function(e) FALSE)
  if (!ok) {
    stop("Database connection is no longer valid. Please re-run the db-connect chunk.")
  }
  invisible(conn)
}

get_complication_descendants <- function(conn, complication_concepts) {

  ensure_connection(conn)

  omop_database <- params$omop_database
  omop_schema   <- params$omop_schema

  ancestor_table <- paste0(
    omop_database, ".", omop_schema, ".concept_ancestor"
  )

  out_list <- lapply(seq_len(nrow(complication_concepts)), function(i) {
    ancestor_id <- complication_concepts$concept_id[i]
    short_name  <- complication_concepts$short_name[i]

    sql <- paste0(
      "SELECT DISTINCT ",
      ancestor_id, " AS ancestor_concept_id, ",
      "'", short_name, "' AS short_name, ",
      "descendant_concept_id ",
      "FROM ", ancestor_table, " ",
      "WHERE ancestor_concept_id = ", ancestor_id, " ",
      "UNION ",
      "SELECT DISTINCT ",
      ancestor_id, " AS ancestor_concept_id, ",
      "'", short_name, "' AS short_name, ",
      ancestor_id, " AS descendant_concept_id"
    )

    db_query(conn, sql) %>%
      tibble::as_tibble()
  })

  dplyr::bind_rows(out_list) %>%
    dplyr::distinct()
}

complication_descendants <- get_complication_descendants(
  conn = con,
  complication_concepts = complication_concepts
)

cat("Complication descendant concepts:\n")
cat("- Total descendant concept rows:", nrow(complication_descendants), "\n")
cat("- Unique descendant concept IDs:",
    dplyr::n_distinct(complication_descendants$descendant_concept_id), "\n")

```



```{r pregnancy-complications-query}
# =========================================================
# Step 1B: query condition_occurrence using descendant IDs
# =========================================================

query_complication_conditions <- function(conn, person_ids, complication_descendants, batch_size = 5000) {

  ensure_connection(conn)

  if (is.null(person_ids) || length(person_ids) == 0) {
    return(tibble::tibble())
  }

  omop_database <- params$omop_database
  omop_schema   <- params$omop_schema

  condition_table <- paste0(
    omop_database, ".", omop_schema, ".condition_occurrence"
  )

  person_ids <- unique(person_ids[!is.na(person_ids)])
  concept_ids <- unique(complication_descendants$descendant_concept_id)

  concept_sql <- paste(concept_ids, collapse = ", ")

  person_batches <- split(
    person_ids,
    ceiling(seq_along(person_ids) / batch_size)
  )

  out_list <- lapply(seq_along(person_batches), function(i) {
    batch_ids <- person_batches[[i]]
    person_sql <- paste(batch_ids, collapse = ", ")

    sql <- paste0(
      "SELECT person_id, ",
      "CAST(condition_start_date AS DATE) AS condition_start_date, ",
      "condition_concept_id ",
      "FROM ", condition_table, " ",
      "WHERE condition_concept_id IN (", concept_sql, ") ",
      "  AND person_id IN (", person_sql, ") ",
      "  AND condition_concept_id <> 0"
    )

    cat("Querying batch", i, "of", length(person_batches), "...\n")

    db_query(conn, sql) %>%
      tibble::as_tibble()
  })

  dplyr::bind_rows(out_list) %>%
    dplyr::mutate(
      condition_start_date = as.Date(condition_start_date)
    ) %>%
    dplyr::inner_join(
      complication_descendants,
      by = c("condition_concept_id" = "descendant_concept_id"),
      relationship = "many-to-many"
    ) %>%
    dplyr::distinct(
      person_id,
      condition_start_date,
      condition_concept_id,
      ancestor_concept_id,
      short_name
    )
}

gt_persons <- unique(match_results$gt_all$person_id)
algo_persons <- unique(algo_predictions$person_id)
persons_of_interest <- unique(c(gt_persons, algo_persons))

complication_conditions <- query_complication_conditions(
  conn = con,
  person_ids = persons_of_interest,
  complication_descendants = complication_descendants,
  batch_size = 5000
)

cat("Pregnancy Complication Condition Records:\n")
cat("- Persons in GT:", length(gt_persons), "\n")
cat("- Persons in algorithm predictions:", length(algo_persons), "\n")
cat("- Persons queried in union:", length(persons_of_interest), "\n")
cat("- Total records queried for target concepts + descendants:", nrow(complication_conditions), "\n")
cat("- Unique persons with at least one target complication code:",
    dplyr::n_distinct(complication_conditions$person_id), "\n")
cat("- Unique complication concept IDs returned:",
    dplyr::n_distinct(complication_conditions$condition_concept_id), "\n")
```




```{r pregnancy-complications-derive-episode-flags}
# =========================================================
# Derive episode-level complication flags
# A complication is considered detected if the condition date
# falls within the episode window.
# Uses short_name already returned from the OMOP query.
# =========================================================

flag_episode_complications <- function(episodes,
                                       person_col,
                                       episode_col,
                                       start_col,
                                       end_col,
                                       conditions) {

  ep <- episodes %>%
    dplyr::transmute(
      person_id = .data[[person_col]],
      episode_id = .data[[episode_col]],
      episode_start = as.Date(.data[[start_col]]),
      episode_end   = as.Date(.data[[end_col]])
    ) %>%
    dplyr::mutate(
      episode_start = dplyr::coalesce(episode_start, episode_end),
      episode_end   = dplyr::coalesce(episode_end, episode_start),
      episode_key = paste(person_id, episode_id, sep = "::")
    ) %>%
    dplyr::filter(!is.na(person_id), !is.na(episode_start), !is.na(episode_end))

  if (nrow(ep) == 0) {
    return(tibble::tibble())
  }

  cond <- conditions %>%
    dplyr::filter(
      !is.na(person_id),
      !is.na(condition_start_date),
      !is.na(short_name)
    ) %>%
    dplyr::distinct(person_id, condition_start_date, condition_concept_id, short_name)

  if (nrow(cond) == 0) {
    return(
      ep %>%
        dplyr::transmute(
          episode_key,
          preeclampsia_flag = 0L,
          pih_flag = 0L,
          gdm_flag = 0L,
          any_complication_flag = 0L
        )
    )
  }

  episode_hits_long <- ep %>%
    dplyr::inner_join(
      cond,
      by = "person_id",
      relationship = "many-to-many"
    ) %>%
    dplyr::filter(
      condition_start_date >= episode_start,
      condition_start_date <= episode_end
    ) %>%
    dplyr::distinct(episode_key, short_name)

  flags_wide <- episode_hits_long %>%
    dplyr::mutate(flag = 1L) %>%
    tidyr::pivot_wider(
      names_from = short_name,
      values_from = flag,
      values_fill = 0
    )

  out <- ep %>%
    dplyr::distinct(episode_key) %>%
    dplyr::left_join(flags_wide, by = "episode_key") %>%
    dplyr::mutate(
      preeclampsia = dplyr::coalesce(preeclampsia, 0L),
      pih = dplyr::coalesce(pih, 0L),
      gdm = dplyr::coalesce(gdm, 0L)
    ) %>%
    dplyr::transmute(
      episode_key,
      preeclampsia_flag = as.integer(preeclampsia > 0),
      pih_flag = as.integer(pih > 0),
      gdm_flag = as.integer(gdm > 0),
      any_complication_flag = as.integer((preeclampsia > 0) | (pih > 0) | (gdm > 0))
    )

  out
}

algo_complication_flags <- flag_episode_complications(
  episodes = algo_predictions,
  person_col = "person_id",
  episode_col = "episode_number",
  start_col = "episode_start_date",
  end_col = "episode_end_date",
  conditions = complication_conditions
)

gt_primary_complication_flags <- flag_episode_complications(
  episodes = match_results$gt_primary,
  person_col = "person_id",
  episode_col = "gt_episode_num",
  start_col = "gt_start",
  end_col = "gt_end",
  conditions = complication_conditions
)

gt_all_complication_flags <- flag_episode_complications(
  episodes = match_results$gt_all,
  person_col = "person_id",
  episode_col = "gt_episode_num",
  start_col = "gt_start",
  end_col = "gt_end",
  conditions = complication_conditions
)

cat("Episode-Level Complication Flags Created:\n")
cat("- Algorithm episodes with flags:", nrow(algo_complication_flags), "\n")
cat("- Eligible GT episodes with flags:", nrow(gt_primary_complication_flags), "\n")
cat("- All GT episodes with flags:", nrow(gt_all_complication_flags), "\n")
```


## Summary

```{r pregnancy-complications-overall-summary}
# =========================================================
# Overall complication prevalence by approach
# =========================================================

summarise_complication_flags <- function(flags, approach_label) {

  flags %>%
    dplyr::summarise(
      preeclampsia = mean(preeclampsia_flag, na.rm = TRUE),
      pih = mean(pih_flag, na.rm = TRUE),
      gdm = mean(gdm_flag, na.rm = TRUE),
      any_complication = mean(any_complication_flag, na.rm = TRUE)
    ) %>%
    dplyr::mutate(approach = approach_label) %>%
    dplyr::select(
      approach,
      preeclampsia,
      pih,
      gdm,
      any_complication
    )
}

complication_prevalence_summary <- dplyr::bind_rows(
  summarise_complication_flags(algo_complication_flags, "Algorithm episodes"),
  summarise_complication_flags(gt_primary_complication_flags, "Eligible GT episodes"),
  summarise_complication_flags(gt_all_complication_flags, "All GT episodes")
)

complication_prevalence_summary %>%
  dplyr::mutate(
    dplyr::across(
      c(preeclampsia, pih, gdm, any_complication),
      ~ round(100 * ., 1)
    )
  ) %>%
  knitr::kable(
    caption = "Overall Pregnancy Complication Prevalence by Approach",
    col.names = c(
      "Approach",
      "Preeclampsia (%)",
      "Pregnancy-Induced Hypertension (%)",
      "Gestational Diabetes (%)",
      "Any Complication (%)"
    )
  ) %>%
  kableExtra::kable_styling(
    bootstrap_options = c("striped", "hover"),
    full_width = FALSE
  )
```


## matched episodes

```{r pregnancy-complications-matched-episodes}
# =========================================================
# Compare complication detection between approaches
# among primary matched episodes
# =========================================================

matched_complications <- matched_episodes %>%
  dplyr::left_join(
    algo_complication_flags %>%
      dplyr::rename(
        algo_preeclampsia_flag = preeclampsia_flag,
        algo_pih_flag = pih_flag,
        algo_gdm_flag = gdm_flag,
        algo_any_complication_flag = any_complication_flag
      ),
    by = c("algo_key" = "episode_key")
  ) %>%
  dplyr::left_join(
    gt_primary_complication_flags %>%
      dplyr::rename(
        gt_preeclampsia_flag = preeclampsia_flag,
        gt_pih_flag = pih_flag,
        gt_gdm_flag = gdm_flag,
        gt_any_complication_flag = any_complication_flag
      ),
    by = c("gt_key" = "episode_key")
  ) %>%
  dplyr::mutate(
    dplyr::across(
      c(
        algo_preeclampsia_flag, algo_pih_flag, algo_gdm_flag, algo_any_complication_flag,
        gt_preeclampsia_flag, gt_pih_flag, gt_gdm_flag, gt_any_complication_flag
      ),
      ~ dplyr::coalesce(., 0L)
    )
  )

cat("Matched Episode Complication Comparison:\n")
cat("- Primary matched episodes available for complication comparison:",
    nrow(matched_complications), "\n")
```


## Agreement summary
```{r pregnancy-complications-agreement-summary}
# =========================================================
# Agreement metrics for complication detection
# GT episode window used as reference
# =========================================================

calculate_binary_comparison_metrics <- function(dat, algo_col, gt_col, label) {

  algo_vec <- dat[[algo_col]]
  gt_vec   <- dat[[gt_col]]

  tp <- sum(algo_vec == 1 & gt_vec == 1, na.rm = TRUE)
  fp <- sum(algo_vec == 1 & gt_vec == 0, na.rm = TRUE)
  fn <- sum(algo_vec == 0 & gt_vec == 1, na.rm = TRUE)
  tn <- sum(algo_vec == 0 & gt_vec == 0, na.rm = TRUE)

  n <- tp + fp + fn + tn

  sensitivity <- ifelse(tp + fn > 0, tp / (tp + fn), NA_real_)
  specificity <- ifelse(tn + fp > 0, tn / (tn + fp), NA_real_)
  ppv <- ifelse(tp + fp > 0, tp / (tp + fp), NA_real_)
  agreement <- ifelse(n > 0, (tp + tn) / n, NA_real_)

  # Cohen's kappa
  p_obs <- agreement
  p_exp <- ifelse(
    n > 0,
    ((tp + fp) / n) * ((tp + fn) / n) +
    ((fn + tn) / n) * ((fp + tn) / n),
    NA_real_
  )

  kappa <- ifelse(
    !is.na(p_exp) && (1 - p_exp) > 0,
    (p_obs - p_exp) / (1 - p_exp),
    NA_real_
  )

  tibble::tibble(
    complication = label,
    n = n,
    algorithm_positive = sum(algo_vec == 1, na.rm = TRUE),
    gt_positive = sum(gt_vec == 1, na.rm = TRUE),
    both_positive = tp,
    algorithm_only = fp,
    gt_only = fn,
    both_negative = tn,
    sensitivity = sensitivity,
    specificity = specificity,
    ppv = ppv,
    agreement = agreement,
    kappa = kappa
  )
}

complication_agreement_summary <- dplyr::bind_rows(
  calculate_binary_comparison_metrics(
    matched_complications,
    "algo_preeclampsia_flag",
    "gt_preeclampsia_flag",
    "Preeclampsia"
  ),
  calculate_binary_comparison_metrics(
    matched_complications,
    "algo_pih_flag",
    "gt_pih_flag",
    "Pregnancy-induced hypertension"
  ),
  calculate_binary_comparison_metrics(
    matched_complications,
    "algo_gdm_flag",
    "gt_gdm_flag",
    "Gestational diabetes"
  ),
  calculate_binary_comparison_metrics(
    matched_complications,
    "algo_any_complication_flag",
    "gt_any_complication_flag",
    "Any complication"
  )
)

complication_agreement_summary %>%
  dplyr::mutate(
    dplyr::across(
      c(sensitivity, specificity, ppv, agreement),
      ~ round(100 * ., 1)
      ),
    kappa = round(kappa, 3)
  ) %>%
  knitr::kable(
    caption = "Pregnancy Complication Detection Agreement Among Primary Matched Episodes",
    col.names = c(
      "Complication",
      "N",
      "Algorithm Positive",
      "GT Positive",
      "Both Positive",
      "Algorithm Only",
      "GT Only",
      "Both Negative",
      "Sensitivity (%)",
      "Specificity (%)",
      "PPV (%)",
      "Agreement (%)",
      "Cohen's Kappa"
)
  ) %>%
  kableExtra::kable_styling(
    bootstrap_options = c("striped", "hover"),
    full_width = FALSE
  )
```

## Detection Rate

```{r pregnancy-complications-detection-rates}
# =========================================================
# Detection rates by approach among matched episodes
# =========================================================

matched_complication_rates <- tibble::tibble(
  complication = c(
    "Preeclampsia",
    "Pregnancy-induced hypertension",
    "Gestational diabetes",
    "Any complication"
  ),
  algorithm_rate = c(
    mean(matched_complications$algo_preeclampsia_flag, na.rm = TRUE),
    mean(matched_complications$algo_pih_flag, na.rm = TRUE),
    mean(matched_complications$algo_gdm_flag, na.rm = TRUE),
    mean(matched_complications$algo_any_complication_flag, na.rm = TRUE)
  ),
  gt_rate = c(
    mean(matched_complications$gt_preeclampsia_flag, na.rm = TRUE),
    mean(matched_complications$gt_pih_flag, na.rm = TRUE),
    mean(matched_complications$gt_gdm_flag, na.rm = TRUE),
    mean(matched_complications$gt_any_complication_flag, na.rm = TRUE)
  )
)

matched_complication_rates %>%
  dplyr::mutate(
    algorithm_rate = round(100 * algorithm_rate, 1),
    gt_rate = round(100 * gt_rate, 1)
  ) %>%
  knitr::kable(
    caption = "Pregnancy Complication Detection Rates Among Primary Matched Episodes",
    col.names = c(
      "Complication",
      "Algorithm Detection Rate (%)",
      "GT Detection Rate (%)"
    )
  ) %>%
  kableExtra::kable_styling(
    bootstrap_options = c("striped", "hover"),
    full_width = FALSE
  )
```


## Plot
```{r pregnancy-complications-plot}
if (nrow(matched_complication_rates) > 0) {

  complication_rates_long <- matched_complication_rates %>%
    tidyr::pivot_longer(
      cols = c(algorithm_rate, gt_rate),
      names_to = "approach",
      values_to = "rate"
    ) %>%
    dplyr::mutate(
      approach = dplyr::recode(
        approach,
        algorithm_rate = "Algorithm window",
        gt_rate = "GT window"
      )
    )

  ggplot2::ggplot(
    complication_rates_long,
    ggplot2::aes(x = complication, y = rate, fill = approach)
  ) +
    ggplot2::geom_col(position = "dodge") +
    ggplot2::scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
    ggplot2::labs(
      title = "Pregnancy Complication Detection Rates Among Primary Matched Episodes",
      x = "Complication",
      y = "Detection Rate",
      fill = "Approach"
    ) +
    ggplot2::theme_minimal() +
    ggplot2::theme(
      axis.text.x = ggplot2::element_text(angle = 20, hjust = 1)
    )
}
```

## Key findings
```{r pregnancy-complications-key-findings}
cat("Pregnancy Complication Findings:\n")
cat(
  "1. Episode-level complication flags were derived for preeclampsia, pregnancy-induced hypertension, and gestational diabetes using direct condition concept IDs within each episode window.\n"
)

cat(
  "2. Among primary matched episodes, any-complication detection was",
  round(100 * mean(matched_complications$algo_any_complication_flag, na.rm = TRUE), 1),
  "% using the algorithm-defined window and",
  round(100 * mean(matched_complications$gt_any_complication_flag, na.rm = TRUE), 1),
  "% using the ground-truth-defined window.\n"
)

cat(
  "3. Complication-specific agreement metrics are reported for preeclampsia, pregnancy-induced hypertension, and gestational diabetes to assess whether the two episode definitions capture similar clinically relevant pregnancy events.\n"
)

cat(
  "4. This section complements the primary outcome-agnostic episode matching analysis by evaluating whether clinically important pregnancy complications are detected within the episodes identified by either approach.\n"
)
```

```{r cleanup2}
#| include: false

# Disconnect from database
if (exists("con")) {
  db_disconnect(con)
  cat("Database connection closed\n")
}
```


