---
title: "Pregnancy Algorithm Validation Report"
subtitle: "Comparison of Algorithm Predictions vs PMAP OB Ground Truth"
author: "OMOP Pregnancy Team"
date: "`r Sys.Date()`"
format:
  html:
    toc: true
    toc-depth: 3
    toc-location: left
    code-fold: true
    code-summary: "Show code"
    theme: flatly
    self-contained: true
execute:
  echo: true
  warning: false
  message: false
params:
  prediction_file: "../output/pregnancy_episodes_2026-03-08.csv"
  date_window_days: 30
  min_ga_days: 0
  max_ga_days: 320
---

```{r setup}
#| label: setup
#| include: false

# Load required packages
library(dplyr)
library(tidyr)
library(purrr)
library(lubridate)
library(ggplot2)
library(scales)
library(knitr)
library(kableExtra)
library(janitor)
Sys.getenv("JAVA_HOME")
library(rJava)
library(dotenv)
dotenv::load_dot_env()
# Source the connection function from the project
source("../R/00_connection/create_connection.R")

# Set ggplot theme
theme_set(theme_minimal(base_size = 12))
```

# Executive Summary

This report validates the pregnancy identification algorithm by comparing its predictions against ground truth data from the PMAP OB registry. The validation covers:

- **Episode-level matching**: Matching predicted episodes to ground truth using a ±`r params$date_window_days` day window
- **Outcome classification**: Accuracy of outcome category predictions (LB, SB, ECT, AB, SA, DELIV, PREG)
- **Gestational age estimation**: Comparison of predicted vs ground truth gestational age
- **Person-level aggregation**: Total pregnancy counts per person

```{r load-predictions}
#| label: load-predictions

# Load algorithm predictions
load_algorithm_predictions <- function(file_path) {
  stopifnot(is.character(file_path), length(file_path) == 1L, nzchar(file_path))
  if (!file.exists(file_path)) stop("File not found: ", file_path)

  ext <- tolower(tools::file_ext(file_path))

  predictions <- if (ext == "rds") {
    readRDS(file_path)
  } else if (ext %in% c("csv", "txt")) {
    read.csv(file_path, stringsAsFactors = FALSE, check.names = FALSE)
  } else {
    stop("Unsupported file extension: .", ext, " (supported: rds, csv, txt)")
  }

  # Require deps explicitly (fail fast with clear message)
  if (!requireNamespace("dplyr", quietly = TRUE)) stop("Package 'dplyr' is required.")
  if (!requireNamespace("janitor", quietly = TRUE)) stop("Package 'janitor' is required.")

  # Helper: robust Date coercion (supports Date/POSIXct/character/numeric)
  to_date_safe <- function(x) {
    if (inherits(x, "Date")) return(x)
    if (inherits(x, "POSIXt")) return(as.Date(x))
    if (is.numeric(x)) return(as.Date(x, origin = "1970-01-01"))
    suppressWarnings(as.Date(x))
  }

  predictions <- predictions %>%
    janitor::clean_names()

  # If *_x columns exist (merge artifacts), coalesce into canonical fields
  if (!"episode_start_date" %in% names(predictions) && "episode_start_date_x" %in% names(predictions)) {
    predictions$episode_start_date <- predictions$episode_start_date_x
  }
  if (!"episode_end_date" %in% names(predictions) && "episode_end_date_x" %in% names(predictions)) {
    predictions$episode_end_date <- predictions$episode_end_date_x
  }

  # Validate required columns
  required_cols <- c(
    "person_id", "episode_number", "episode_start_date", "episode_end_date",
    "outcome_category", "gestational_age_days", "algorithm_used", "precision_category"
  )
  missing <- setdiff(required_cols, names(predictions))
  if (length(missing) > 0) {
    stop("Missing required columns after clean_names(): ", paste(missing, collapse = ", "))
  }

  # Clean / type-cast and select canonical fields
  predictions <- predictions %>%
    dplyr::mutate(
      person_id = suppressWarnings(as.integer(person_id)),
      episode_number = suppressWarnings(as.integer(episode_number)),
      episode_start_date = to_date_safe(episode_start_date),
      episode_end_date = to_date_safe(episode_end_date),
      gestational_age_days = suppressWarnings(as.integer(gestational_age_days)),
      outcome_category = as.character(outcome_category),
      algorithm_used = as.character(algorithm_used),
      precision_category = as.character(precision_category)
    ) %>%
    dplyr::select(dplyr::all_of(required_cols)) %>%
    dplyr::distinct()

  return(predictions)
}

# Load predictions
algo_predictions <- load_algorithm_predictions(params$prediction_file)

cat(
  "Loaded", nrow(algo_predictions), "algorithm predictions for",
  dplyr::n_distinct(algo_predictions$person_id), "unique persons\n"
)
```

```{r db-connect}
#| label: db-connect
# Connect to database using project connection function
con <- create_connection_from_env(".env")
if (is.null(con)) stop("Connection returned NULL.")
if (!DBI::dbIsValid(con)) stop("Database connection is not valid.")

cat("Database connection established\n")
```

```{r query-ground-truth}
#| label: query-ground-truth

# Query ground truth tables

# 1. Pregnancy episodes
pregnancies_sql <- "
SELECT
  EMRN,
  PregnancyEstimatedStartDate,
  PregnancyEstimatedEndDate,
  PregnancyGravidaCount,
  PregnancyParaCount,
  PregnancyAbortionCount,
  PregnancyTherapeuticAbortionCount,
  PregnancySpontaneousAbortionCount,
  PregnancyEctopicCount
FROM Obstetrics_Minhas_IRB00501137_Scratch.phi.PMAP_OB_PREGNANCIES
"

gt_pregnancies <- DatabaseConnector::querySql(con, pregnancies_sql) %>%
  clean_names() %>%
  mutate(
    pregnancy_estimated_start_date = as.Date(pregnancy_estimated_start_date),
    pregnancy_estimated_end_date = as.Date(pregnancy_estimated_end_date)
  )

cat("Loaded", nrow(gt_pregnancies), "ground truth pregnancy episodes\n")
cat("Sample EMRNs from pregnancies:", head(gt_pregnancies$emrn, 5), "\n")

# 2. Delivery outcomes
deliveries_sql <- "
SELECT
  mom_emrn AS emrn,
  GestationalAgeDays,
  LivingStatus,
  BornAlive,
  NeonatalDemise
FROM Obstetrics_Minhas_IRB00501137_Scratch.phi.PMAP_OB_DEL
"

gt_deliveries <- DatabaseConnector::querySql(con, deliveries_sql) %>%
  clean_names()

cat("Loaded", nrow(gt_deliveries), "ground truth delivery records\n")

# 3. EMRN to cohort_id linkage
emrn_cohort_sql <- "
SELECT emrn, cohort_id
FROM Obstetrics_Minhas_IRB00501137_Projection.phi.derived_epic_patient
"

emrn_to_cohort <- DatabaseConnector::querySql(con, emrn_cohort_sql) %>%
  clean_names()

cat("Loaded", nrow(emrn_to_cohort), "EMRN to cohort_id mappings\n")
cat("Sample EMRNs from linkage:", head(emrn_to_cohort$emrn, 5), "\n")

# Check EMRN overlap between pregnancies and linkage
preg_emrns <- unique(gt_pregnancies$emrn)
link_emrns <- unique(emrn_to_cohort$emrn)
emrn_overlap <- length(intersect(preg_emrns, link_emrns))
cat("EMRN overlap (pregnancies ∩ linkage):", emrn_overlap, "/", length(preg_emrns), "\n")

# 4. cohort_id (pmap_id) to person_id linkage
cohort_person_sql <- "
SELECT pmap_id AS cohort_id, person_id
FROM Obstetrics_Minhas_IRB00501137_OMOP.dbo.registry_idmap
"

cohort_to_person <- DatabaseConnector::querySql(con, cohort_person_sql) %>%
  clean_names()

cat("Loaded", nrow(cohort_to_person), "cohort_id to person_id mappings\n")
```

```{r id-linkage}
#| label: id-linkage

# Build complete ID linkage: EMRN -> cohort_id -> person_id
build_id_linkage <- function(emrn_cohort, cohort_person) {
  linkage <- emrn_cohort %>%
    inner_join(cohort_person, by = "cohort_id") %>%
    select(emrn, person_id) %>%
    distinct()

  return(linkage)
}

# Debug: Check cohort_id overlap before building linkage
cohort_ids_from_emrn <- unique(emrn_to_cohort$cohort_id)
cohort_ids_from_person <- unique(cohort_to_person$cohort_id)
cohort_overlap <- length(intersect(cohort_ids_from_emrn, cohort_ids_from_person))
cat("Cohort_id overlap (emrn_table ∩ person_table):", cohort_overlap, "\n")
cat("Sample cohort_ids from emrn table:", head(cohort_ids_from_emrn, 5), "\n")
cat("Sample cohort_ids from person table:", head(cohort_ids_from_person, 5), "\n")

id_linkage <- build_id_linkage(emrn_to_cohort, cohort_to_person)

cat("\nCreated ID linkage table with", nrow(id_linkage), "unique EMRN-person_id pairs\n")
cat("Columns in id_linkage:", names(id_linkage), "\n")

# Debug: Show sample of linkage
cat("Sample person_ids in linkage:", head(id_linkage$person_id, 10), "\n")
cat("Sample emrns in linkage:", head(id_linkage$emrn, 5), "\n")

# Check linkage coverage
algo_persons <- n_distinct(algo_predictions$person_id)
algo_person_ids <- unique(algo_predictions$person_id)
linked_persons <- sum(algo_person_ids %in% id_linkage$person_id)
cat("Algorithm persons with ID linkage:", linked_persons, "/", algo_persons,
    "(", round(100 * linked_persons / algo_persons, 1), "%)\n")

# Debug: Show sample algorithm person_ids
cat("Sample algorithm person_ids:", head(algo_person_ids, 10), "\n")

# Check for any overlap
overlap_count <- length(intersect(algo_person_ids, id_linkage$person_id))
cat("Overlapping person_ids between algorithm and linkage:", overlap_count, "\n")
```

```{r derive-outcomes}
#| label: derive-outcomes

# Derive outcome category from ground truth cumulative counts
derive_ground_truth_outcome <- function(pregnancies_df, deliveries_df, id_linkage) {

  # Ensure count columns are numeric (may come as character from SQL)
  pregnancies_df <- pregnancies_df %>%
    mutate(
      pregnancy_gravida_count = as.numeric(pregnancy_gravida_count),
      pregnancy_para_count = as.numeric(pregnancy_para_count),
      pregnancy_abortion_count = as.numeric(pregnancy_abortion_count),
      pregnancy_therapeutic_abortion_count = as.numeric(pregnancy_therapeutic_abortion_count),
      pregnancy_spontaneous_abortion_count = as.numeric(pregnancy_spontaneous_abortion_count),
      pregnancy_ectopic_count = as.numeric(pregnancy_ectopic_count)
    )

  # Join pregnancies with ID linkage to get person_id
  pregnancies_linked <- pregnancies_df %>%
    inner_join(id_linkage, by = "emrn")

  # Ensure delivery columns are proper types
  deliveries_df <- deliveries_df %>%
    mutate(
      gestational_age_days = as.numeric(gestational_age_days),
      born_alive = as.numeric(born_alive)
    )

  # Join with deliveries (if available) for BornAlive info
  # Deliveries may have multiple records per EMRN, take the one closest to pregnancy end
  deliveries_by_emrn <- deliveries_df %>%
    group_by(emrn) %>%
    summarise(
      gestational_age_days = first(na.omit(gestational_age_days)),
      born_alive = first(na.omit(born_alive)),
      living_status = first(na.omit(living_status)),
      .groups = "drop"
    )

  # Combine and derive outcomes
  derived <- pregnancies_linked %>%
    left_join(deliveries_by_emrn, by = "emrn") %>%
    group_by(person_id) %>%
    arrange(pregnancy_gravida_count) %>%
    mutate(
      # Calculate deltas from previous episode
      prev_para = lag(pregnancy_para_count, default = 0),
      prev_ab = lag(pregnancy_abortion_count, default = 0),
      prev_therapeutic = lag(pregnancy_therapeutic_abortion_count, default = 0),
      prev_spontaneous = lag(pregnancy_spontaneous_abortion_count, default = 0),
      prev_ectopic = lag(pregnancy_ectopic_count, default = 0),

      delta_para = pregnancy_para_count - prev_para,
      delta_ab = pregnancy_abortion_count - prev_ab,
      delta_therapeutic = pregnancy_therapeutic_abortion_count - prev_therapeutic,
      delta_spontaneous = pregnancy_spontaneous_abortion_count - prev_spontaneous,
      delta_ectopic = pregnancy_ectopic_count - prev_ectopic,

      # Derive outcome category following Matcho hierarchy
      gt_outcome_category = case_when(
        # Live birth: para increased and born alive
        delta_para > 0 & !is.na(born_alive) & born_alive == 1 ~ "LB",
        # Stillbirth: para increased and not born alive (or living_status indicates)
        delta_para > 0 & (!is.na(born_alive) & born_alive == 0) ~ "SB",
        delta_para > 0 & (!is.na(living_status) & tolower(living_status) %in% c("stillborn", "fetal demise")) ~ "SB",
        # Ectopic pregnancy
        delta_ectopic > 0 ~ "ECT",
        # Therapeutic/Induced abortion
        delta_therapeutic > 0 ~ "AB",
        # Spontaneous abortion
        delta_spontaneous > 0 ~ "SA",
        # Generic abortion (when not specified as therapeutic or spontaneous)
        delta_ab > 0 & delta_therapeutic == 0 & delta_spontaneous == 0 ~ "AB",
        # Delivery without specific birth info
        delta_para > 0 ~ "DELIV",
        # Pregnancy without recorded outcome
        TRUE ~ "PREG"
      )
    ) %>%
    ungroup() %>%
    select(
      person_id,
      emrn,
      gt_episode_num = pregnancy_gravida_count,
      gt_start = pregnancy_estimated_start_date,
      gt_end = pregnancy_estimated_end_date,
      gt_outcome_category,
      gt_ga_days = gestational_age_days,
      born_alive,
      living_status
    )

  return(derived)
}

gt_episodes <- derive_ground_truth_outcome(gt_pregnancies, gt_deliveries, id_linkage)

cat("Derived outcomes for", nrow(gt_episodes), "ground truth episodes\n")
cat("Unique persons in ground truth:", n_distinct(gt_episodes$person_id), "\n")

# Debug: Check overlap between algorithm and ground truth persons
algo_persons_set <- unique(algo_predictions$person_id)
gt_persons_set <- unique(gt_episodes$person_id)
person_overlap <- length(intersect(algo_persons_set, gt_persons_set))
cat("Person overlap (algo ∩ GT):", person_overlap, "\n")

# Debug: Sample dates
cat("\nSample algorithm end dates:", head(algo_predictions$episode_end_date, 5), "\n")
cat("Sample GT end dates:", head(gt_episodes$gt_end, 5), "\n")

cat("\nGround truth outcome distribution:\n")
gt_episodes %>%
  count(gt_outcome_category) %>%
  arrange(desc(n)) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
```

# Episode-level matching
```{r episode-level-matching}
# =========================================================
# Episode-level matching
# Primary analysis: overlap-based one-to-one matching
# Sensitivity analyses: end-date windows (±14 / ±30 / ±60 days)
# =========================================================

# Prepare episode inputs for validation
prepare_validation_inputs <- function(algo_episodes, gt_episodes) {

  algo <- algo_episodes %>%
    dplyr::select(
      person_id,
      algo_episode_num = episode_number,
      algo_start = episode_start_date,
      algo_end = episode_end_date,
      algo_outcome = outcome_category,
      algo_ga_days = gestational_age_days,
      algo_method = dplyr::any_of("algorithm_used")
    ) %>%
    dplyr::mutate(
      algo_start = as.Date(algo_start),
      algo_end   = as.Date(algo_end)
    ) %>%
    dplyr::filter(!is.na(person_id), !is.na(algo_end)) %>%
    dplyr::group_by(person_id) %>%
    dplyr::arrange(algo_end, .by_group = TRUE) %>%
    dplyr::mutate(algo_order = dplyr::row_number()) %>%
    dplyr::ungroup()

  gt <- gt_episodes %>%
    dplyr::select(
      person_id,
      gt_episode_num,
      gt_start,
      gt_end,
      gt_outcome = gt_outcome_category,
      gt_ga_days
    ) %>%
    dplyr::mutate(
      gt_start = as.Date(gt_start),
      gt_end   = as.Date(gt_end)
    ) %>%
    dplyr::filter(!is.na(person_id), !is.na(gt_end)) %>%
    dplyr::group_by(person_id) %>%
    dplyr::arrange(gt_end, .by_group = TRUE) %>%
    dplyr::mutate(gt_order = dplyr::row_number()) %>%
    dplyr::ungroup()

  list(algo = algo, gt = gt)
}

# Create candidate pairs within each person
build_match_candidates <- function(algo, gt) {

  common_persons <- intersect(unique(algo$person_id), unique(gt$person_id))

  cat("Persons in algorithm:", length(unique(algo$person_id)), "\n")
  cat("Persons in ground truth:", length(unique(gt$person_id)), "\n")
  cat("Person overlap for matching:", length(common_persons), "\n")

  if (length(common_persons) == 0) {
    warning("No person_id overlap between algorithm and ground truth.")
    return(tibble::tibble())
  }

  algo_sub <- algo %>%
    dplyr::filter(person_id %in% common_persons) %>%
    dplyr::mutate(.join_key = 1L)

  gt_sub <- gt %>%
    dplyr::filter(person_id %in% common_persons) %>%
    dplyr::mutate(.join_key = 1L)

  # Intentional many-to-many join:
  # within each person_id, generate all possible algorithm-ground truth episode pairs
  candidates <- algo_sub %>%
    dplyr::inner_join(
      gt_sub,
      by = c("person_id", ".join_key"),
      relationship = "many-to-many"
    ) %>%
    dplyr::select(-.join_key) %>%
    dplyr::mutate(
      overlap_start = pmax(algo_start, gt_start, na.rm = TRUE),
      overlap_end   = pmin(algo_end, gt_end, na.rm = TRUE),
      overlap_days  = pmax(0, as.numeric(overlap_end - overlap_start) + 1),
      overlap_any   = overlap_days > 0,
      start_diff_days = abs(as.numeric(algo_start - gt_start)),
      end_diff_days   = abs(as.numeric(algo_end - gt_end)),
      mid_algo = algo_start + floor(as.numeric(algo_end - algo_start) / 2),
      mid_gt   = gt_start + floor(as.numeric(gt_end - gt_start) / 2),
      midpoint_diff_days = abs(as.numeric(mid_algo - mid_gt))
    ) %>%
    dplyr::mutate(
      algo_key = paste(person_id, algo_episode_num, sep = "::"),
      gt_key   = paste(person_id, gt_episode_num, sep = "::")
    )

  return(candidates)
}

# Greedy one-to-one matcher
# rule = "overlap" or "window"
greedy_one_to_one_match <- function(candidates,
                                    algo,
                                    gt,
                                    rule = c("overlap", "window"),
                                    date_window_days = 30) {

  rule <- match.arg(rule)

  if (nrow(candidates) == 0) {
    return(list(
      matched = tibble::tibble(),
      unmatched_algo = algo,
      unmatched_gt = gt,
      stats = list(
        total_algo = nrow(algo),
        total_gt = nrow(gt),
        matched = 0L,
        match_rate_algo = 0,
        match_rate_gt = 0
      )
    ))
  }

  candidates_ranked <- candidates

  if (rule == "overlap") {
    candidates_ranked <- candidates_ranked %>%
      dplyr::filter(overlap_any) %>%
      dplyr::arrange(
        person_id,
        dplyr::desc(overlap_days),
        end_diff_days,
        start_diff_days,
        midpoint_diff_days,
        algo_order,
        gt_order
      )
  }

  if (rule == "window") {
    candidates_ranked <- candidates_ranked %>%
      dplyr::filter(end_diff_days <= date_window_days) %>%
      dplyr::arrange(
        person_id,
        end_diff_days,
        start_diff_days,
        midpoint_diff_days,
        dplyr::desc(overlap_days),
        algo_order,
        gt_order
      )
  }

  if (nrow(candidates_ranked) == 0) {
    return(list(
      matched = tibble::tibble(),
      unmatched_algo = algo,
      unmatched_gt = gt,
      stats = list(
        total_algo = nrow(algo),
        total_gt = nrow(gt),
        matched = 0L,
        match_rate_algo = 0,
        match_rate_gt = 0
      )
    ))
  }

  used_algo <- character(0)
  used_gt <- character(0)
  keep <- logical(nrow(candidates_ranked))

  for (i in seq_len(nrow(candidates_ranked))) {
    this_algo <- candidates_ranked$algo_key[i]
    this_gt   <- candidates_ranked$gt_key[i]

    if (!(this_algo %in% used_algo) && !(this_gt %in% used_gt)) {
      keep[i] <- TRUE
      used_algo <- c(used_algo, this_algo)
      used_gt   <- c(used_gt, this_gt)
    }
  }

  matched <- candidates_ranked[keep, , drop = FALSE]

  unmatched_algo <- algo %>%
    dplyr::filter(!(paste(person_id, algo_episode_num, sep = "::") %in% matched$algo_key))

  unmatched_gt <- gt %>%
    dplyr::filter(!(paste(person_id, gt_episode_num, sep = "::") %in% matched$gt_key))

  stats <- list(
    total_algo = nrow(algo),
    total_gt = nrow(gt),
    matched = nrow(matched),
    match_rate_algo = ifelse(nrow(algo) > 0, nrow(matched) / nrow(algo), 0),
    match_rate_gt   = ifelse(nrow(gt) > 0, nrow(matched) / nrow(gt), 0)
  )

  list(
    matched = matched,
    unmatched_algo = unmatched_algo,
    unmatched_gt = unmatched_gt,
    stats = stats
  )
}

# Wrapper to run all matching analyses
run_matching_analyses <- function(algo_episodes, gt_episodes) {

  prepared <- prepare_validation_inputs(algo_episodes, gt_episodes)
  algo <- prepared$algo
  gt   <- prepared$gt

  candidates <- build_match_candidates(algo, gt)

  primary_overlap <- greedy_one_to_one_match(
    candidates = candidates,
    algo = algo,
    gt = gt,
    rule = "overlap"
  )

  end_window_14 <- greedy_one_to_one_match(
    candidates = candidates,
    algo = algo,
    gt = gt,
    rule = "window",
    date_window_days = 14
  )

  end_window_30 <- greedy_one_to_one_match(
    candidates = candidates,
    algo = algo,
    gt = gt,
    rule = "window",
    date_window_days = 30
  )

  end_window_60 <- greedy_one_to_one_match(
    candidates = candidates,
    algo = algo,
    gt = gt,
    rule = "window",
    date_window_days = 60
  )

  summary_table <- tibble::tibble(
    matching_rule = c(
      "Primary: any episode overlap",
      "Sensitivity: end date within ±14 days",
      "Sensitivity: end date within ±30 days",
      "Sensitivity: end date within ±60 days"
    ),
    matched_episodes = c(
      primary_overlap$stats$matched,
      end_window_14$stats$matched,
      end_window_30$stats$matched,
      end_window_60$stats$matched
    ),
    algorithm_match_rate = c(
      primary_overlap$stats$match_rate_algo,
      end_window_14$stats$match_rate_algo,
      end_window_30$stats$match_rate_algo,
      end_window_60$stats$match_rate_algo
    ),
    ground_truth_match_rate = c(
      primary_overlap$stats$match_rate_gt,
      end_window_14$stats$match_rate_gt,
      end_window_30$stats$match_rate_gt,
      end_window_60$stats$match_rate_gt
    )
  )

  list(
    algo = algo,
    gt = gt,
    candidates = candidates,
    primary_overlap = primary_overlap,
    end_window_14 = end_window_14,
    end_window_30 = end_window_30,
    end_window_60 = end_window_60,
    summary_table = summary_table
  )
}

match_results <- run_matching_analyses(
  algo_episodes = algo_predictions,
  gt_episodes   = gt_episodes
)

matched_episodes <- match_results$primary_overlap$matched

cat("Episode Matching Results (primary overlap-based analysis):\n")
cat("- Algorithm episodes:", match_results$primary_overlap$stats$total_algo, "\n")
cat("- Ground truth episodes:", match_results$primary_overlap$stats$total_gt, "\n")
cat("- Matched episodes:", match_results$primary_overlap$stats$matched, "\n")
cat(
  "- Algorithm match rate:",
  round(100 * match_results$primary_overlap$stats$match_rate_algo, 1),
  "%\n"
)
cat(
  "- Ground truth match rate:",
  round(100 * match_results$primary_overlap$stats$match_rate_gt, 1),
  "%\n"
)
```

## Matching summary table
```{r matching-summary-table}
# Matching summary table
match_results$summary_table %>%
  dplyr::mutate(
    algorithm_match_rate = round(100 * algorithm_match_rate, 1),
    ground_truth_match_rate = round(100 * ground_truth_match_rate, 1)
  ) %>%
  knitr::kable(
    caption = "Episode Matching Summary Across Matching Rules",
    col.names = c(
      "Matching Rule",
      "Matched Episodes",
      "Algorithm Match Rate (%)",
      "Ground Truth Match Rate (%)"
    )
  ) %>%
  kableExtra::kable_styling(
    bootstrap_options = c("striped", "hover"),
    full_width = FALSE
  )
```



# Outcome classification analysis
```{r outcome-classification}
# =========================================================
# Outcome classification analysis
# Secondary analysis: evaluate on primary matched episodes
# =========================================================

create_confusion_matrix <- function(matched_episodes) {

  outcome_order <- c("LB", "SB", "ECT", "AB", "SA", "DELIV", "PREG")

  conf_matrix <- matched_episodes %>%
    dplyr::mutate(
      algo_outcome = factor(algo_outcome, levels = outcome_order),
      gt_outcome   = factor(gt_outcome, levels = outcome_order)
    ) %>%
    dplyr::count(gt_outcome, algo_outcome, .drop = FALSE) %>%
    tidyr::pivot_wider(
      names_from = algo_outcome,
      values_from = n,
      values_fill = 0
    ) %>%
    dplyr::arrange(gt_outcome)

  conf_matrix
}

calculate_classification_metrics <- function(matched_episodes) {

  categories <- sort(unique(c(
    as.character(stats::na.omit(matched_episodes$algo_outcome)),
    as.character(stats::na.omit(matched_episodes$gt_outcome))
  )))

  if (length(categories) == 0 || nrow(matched_episodes) == 0) {
    return(list(
      overall_accuracy = NA_real_,
      total_correct = 0L,
      total_matched = 0L,
      by_category = tibble::tibble()
    ))
  }

  metrics_by_category <- purrr::map_dfr(categories, function(cat) {

    tp <- sum(matched_episodes$algo_outcome == cat & matched_episodes$gt_outcome == cat, na.rm = TRUE)
    fp <- sum(matched_episodes$algo_outcome == cat & matched_episodes$gt_outcome != cat, na.rm = TRUE)
    fn <- sum(matched_episodes$algo_outcome != cat & matched_episodes$gt_outcome == cat, na.rm = TRUE)

    precision <- ifelse(tp + fp > 0, tp / (tp + fp), NA_real_)
    recall    <- ifelse(tp + fn > 0, tp / (tp + fn), NA_real_)
    f1_score  <- ifelse(
      !is.na(precision) && !is.na(recall) && (precision + recall) > 0,
      2 * precision * recall / (precision + recall),
      NA_real_
    )

    tibble::tibble(
      category = cat,
      support = sum(matched_episodes$gt_outcome == cat, na.rm = TRUE),
      predicted = sum(matched_episodes$algo_outcome == cat, na.rm = TRUE),
      tp = tp,
      fp = fp,
      fn = fn,
      precision = precision,
      recall = recall,
      f1_score = f1_score
    )
  })

  total_correct <- sum(matched_episodes$algo_outcome == matched_episodes$gt_outcome, na.rm = TRUE)
  total_matched <- nrow(matched_episodes)
  overall_accuracy <- ifelse(total_matched > 0, total_correct / total_matched, NA_real_)

  list(
    overall_accuracy = overall_accuracy,
    total_correct = total_correct,
    total_matched = total_matched,
    by_category = metrics_by_category
  )
}

conf_matrix <- create_confusion_matrix(matched_episodes)
class_metrics <- calculate_classification_metrics(matched_episodes)

cat("Outcome Classification Results (primary overlap-matched episodes):\n")
cat(
  "- Overall accuracy:",
  round(100 * class_metrics$overall_accuracy, 1),
  "%\n"
)
cat(
  "(",
  class_metrics$total_correct,
  "/",
  class_metrics$total_matched,
  "matched episodes)\n\n"
)
```

## Confusion Matrix
```{r confusion-matrix}
# Confusion matrix
if (nrow(conf_matrix) > 0) {
  conf_matrix %>%
    knitr::kable(caption = "Confusion Matrix: Ground Truth vs Algorithm Outcome") %>%
    kableExtra::kable_styling(
      bootstrap_options = c("striped", "hover"),
      full_width = FALSE
    )
}
```


## Classification metrics by category
```{r classifcation-metrics}
# Classification metrics by category
if (nrow(class_metrics$by_category) > 0) {
  class_metrics$by_category %>%
    dplyr::mutate(
      precision = round(precision, 3),
      recall = round(recall, 3),
      f1_score = round(f1_score, 3)
    ) %>%
    knitr::kable(
      caption = "Classification Metrics by Outcome Category"
    ) %>%
    kableExtra::kable_styling(
      bootstrap_options = c("striped", "hover"),
      full_width = FALSE
    )
}
```

## Plot metrics by category
```{r plot-metrics}
# Plot metrics by category
if (nrow(class_metrics$by_category) > 0) {

  metrics_long <- class_metrics$by_category %>%
    dplyr::filter(support > 0) %>%
    dplyr::select(category, precision, recall, f1_score) %>%
    tidyr::pivot_longer(
      cols = c(precision, recall, f1_score),
      names_to = "metric",
      values_to = "value"
    )

  ggplot2::ggplot(metrics_long, ggplot2::aes(x = category, y = value, fill = metric)) +
    ggplot2::geom_col(position = "dodge") +
    ggplot2::scale_y_continuous(limits = c(0, 1)) +
    ggplot2::labs(
      title = "Outcome Classification Metrics by Category",
      x = "Outcome Category",
      y = "Metric Value",
      fill = "Metric"
    ) +
    ggplot2::theme_minimal()
}
```

# Gestational age comparison

```{r gestational-age}
# =========================================================
# Gestational age comparison
# Primary analysis: overlap-matched episodes
# =========================================================

calculate_ga_metrics <- function(matched_episodes, min_ga = 0, max_ga = 320) {

  with_ga <- matched_episodes %>%
    dplyr::filter(
      !is.na(algo_ga_days),
      !is.na(gt_ga_days),
      algo_ga_days >= min_ga,
      algo_ga_days <= max_ga,
      gt_ga_days >= min_ga,
      gt_ga_days <= max_ga
    ) %>%
    dplyr::mutate(
      ga_diff = algo_ga_days - gt_ga_days,
      abs_ga_diff = abs(ga_diff)
    )

  if (nrow(with_ga) == 0) {
    return(list(
      summary = tibble::tibble(),
      by_outcome = tibble::tibble(),
      data = with_ga
    ))
  }

  summary <- with_ga %>%
    dplyr::summarise(
      n = dplyr::n(),
      mean_diff_days = mean(ga_diff, na.rm = TRUE),
      median_diff_days = stats::median(ga_diff, na.rm = TRUE),
      mean_abs_diff_days = mean(abs_ga_diff, na.rm = TRUE),
      median_abs_diff_days = stats::median(abs_ga_diff, na.rm = TRUE),
      within_7_days = mean(abs_ga_diff <= 7, na.rm = TRUE),
      within_14_days = mean(abs_ga_diff <= 14, na.rm = TRUE),
      within_21_days = mean(abs_ga_diff <= 21, na.rm = TRUE)
    )

  by_outcome <- with_ga %>%
    dplyr::group_by(gt_outcome) %>%
    dplyr::summarise(
      n = dplyr::n(),
      mean_diff_days = mean(ga_diff, na.rm = TRUE),
      median_diff_days = stats::median(ga_diff, na.rm = TRUE),
      mean_abs_diff_days = mean(abs_ga_diff, na.rm = TRUE),
      within_7_days = mean(abs_ga_diff <= 7, na.rm = TRUE),
      within_14_days = mean(abs_ga_diff <= 14, na.rm = TRUE),
      .groups = "drop"
    ) %>%
    dplyr::arrange(dplyr::desc(n))

  list(
    summary = summary,
    by_outcome = by_outcome,
    data = with_ga
  )
}

ga_metrics <- calculate_ga_metrics(matched_episodes)

if (nrow(ga_metrics$summary) > 0) {
  cat("Gestational Age Results (primary overlap-matched episodes with GA available):\n")
  cat("- Mean GA difference:", round(ga_metrics$summary$mean_diff_days, 1), "days\n")
  cat("- Median GA difference:", round(ga_metrics$summary$median_diff_days, 1), "days\n")
  cat("- Mean absolute GA difference:", round(ga_metrics$summary$mean_abs_diff_days, 1), "days\n")
  cat("- Within ±7 days:", round(100 * ga_metrics$summary$within_7_days, 1), "%\n")
  cat("- Within ±14 days:", round(100 * ga_metrics$summary$within_14_days, 1), "%\n")
  cat("- Within ±21 days:", round(100 * ga_metrics$summary$within_21_days, 1), "%\n")
}
```

## GA summary table
```{r GA-summary}
# GA summary table
if (nrow(ga_metrics$summary) > 0) {
  ga_metrics$summary %>%
    dplyr::mutate(dplyr::across(where(is.numeric), ~ round(., 2))) %>%
    tidyr::pivot_longer(
      everything(),
      names_to = "Metric",
      values_to = "Value"
    ) %>%
    knitr::kable(caption = "Gestational Age Comparison Summary") %>%
    kableExtra::kable_styling(
      bootstrap_options = c("striped", "hover"),
      full_width = FALSE
    )
}
```


## GA by outcome category
```{r GA-outcome-category}
# GA by outcome category
if (nrow(ga_metrics$by_outcome) > 0) {
  ga_metrics$by_outcome %>%
    dplyr::mutate(
      dplyr::across(
        where(is.numeric) & !dplyr::matches("^n$"),
        ~ round(., 2)
      )
    ) %>%
    knitr::kable(caption = "GA Agreement by Ground Truth Outcome Category") %>%
    kableExtra::kable_styling(
      bootstrap_options = c("striped", "hover"),
      full_width = FALSE
    )
}
```


## GA agreement plots
```{r GA-agreement}
# GA agreement plots
if (nrow(ga_metrics$data) > 0) {

  ggplot2::ggplot(
    ga_metrics$data,
    ggplot2::aes(x = gt_ga_days, y = algo_ga_days)
  ) +
    ggplot2::geom_point(alpha = 0.3, size = 1.5) +
    ggplot2::geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "red") +
    ggplot2::labs(
      title = "Predicted vs Ground Truth Gestational Age",
      x = "Ground Truth GA (days)",
      y = "Algorithm GA (days)"
    ) +
    ggplot2::theme_minimal()
}
```


## GA metrics plot
```{r GA-metrics-plot}
if (nrow(ga_metrics$data) > 0) {

  ba_data <- ga_metrics$data %>%
    dplyr::mutate(
      mean_ga = (algo_ga_days + gt_ga_days) / 2,
      diff_ga = algo_ga_days - gt_ga_days
    )

  mean_diff <- mean(ba_data$diff_ga, na.rm = TRUE)
  sd_diff <- stats::sd(ba_data$diff_ga, na.rm = TRUE)

  ggplot2::ggplot(
    ba_data,
    ggplot2::aes(x = mean_ga, y = diff_ga)
  ) +
    ggplot2::geom_point(alpha = 0.3, size = 1.5) +
    ggplot2::geom_hline(yintercept = mean_diff, linetype = "solid", color = "red") +
    ggplot2::geom_hline(yintercept = mean_diff + 1.96 * sd_diff, linetype = "dashed", color = "gray40") +
    ggplot2::geom_hline(yintercept = mean_diff - 1.96 * sd_diff, linetype = "dashed", color = "gray40") +
    ggplot2::labs(
      title = "Bland-Altman Plot for Gestational Age",
      x = "Mean GA (days)",
      y = "Algorithm - Ground Truth GA (days)"
    ) +
    ggplot2::theme_minimal()
}
```


# Personal-level aggregation
```{r person-level-aggregation}
# =========================================================
# Person-level aggregation
# Keep exact match, but add within ±1 episode and MAE
# =========================================================

calculate_person_metrics <- function(algo_episodes, gt_episodes) {

  algo_counts <- algo_episodes %>%
    dplyr::group_by(person_id) %>%
    dplyr::summarise(
      algo_total = dplyr::n(),
      .groups = "drop"
    )

  gt_counts <- gt_episodes %>%
    dplyr::group_by(person_id) %>%
    dplyr::summarise(
      gt_total = dplyr::n(),
      .groups = "drop"
    )

  person_data <- dplyr::full_join(algo_counts, gt_counts, by = "person_id") %>%
    dplyr::mutate(
      algo_total = dplyr::coalesce(algo_total, 0L),
      gt_total   = dplyr::coalesce(gt_total, 0L),
      total_diff = algo_total - gt_total,
      abs_total_diff = abs(total_diff),
      exact_match = algo_total == gt_total,
      within_1 = abs_total_diff <= 1
    )

  summary <- person_data %>%
    dplyr::summarise(
      n_persons = dplyr::n(),
      exact_match_rate = mean(exact_match, na.rm = TRUE),
      within_1_rate = mean(within_1, na.rm = TRUE),
      mean_abs_diff = mean(abs_total_diff, na.rm = TRUE),
      median_abs_diff = stats::median(abs_total_diff, na.rm = TRUE)
    )

  list(
    summary = summary,
    person_data = person_data
  )
}

person_metrics <- calculate_person_metrics(
  algo_episodes = algo_predictions,
  gt_episodes   = gt_episodes
)

cat("Person-Level Aggregation Results:\n")
cat(
  "- Exact count match:",
  round(100 * person_metrics$summary$exact_match_rate, 1),
  "%\n"
)
cat(
  "- Within ±1 episode:",
  round(100 * person_metrics$summary$within_1_rate, 1),
  "%\n"
)
cat(
  "- Mean absolute count difference:",
  round(person_metrics$summary$mean_abs_diff, 2),
  "\n"
)
cat(
  "- Median absolute count difference:",
  round(person_metrics$summary$median_abs_diff, 2),
  "\n"
)
```


## Person-level summary table
```{r person-level-summary}
# Person-level summary table
person_metrics$summary %>%
  tidyr::pivot_longer(
    everything(),
    names_to = "Metric",
    values_to = "Value"
  ) %>%
  dplyr::mutate(
    Value = round(Value, 2)
  ) %>%
  knitr::kable(caption = "Person-Level Aggregation Summary") %>%
  kableExtra::kable_styling(
    bootstrap_options = c("striped", "hover"),
    full_width = FALSE
  )
```


## Distribution of count differences
```{r dist-cnt-diff}
# Distribution of count differences
if (nrow(person_metrics$person_data) > 0) {
  ggplot2::ggplot(
    person_metrics$person_data,
    ggplot2::aes(x = total_diff)
  ) +
    ggplot2::geom_histogram(
      binwidth = 1,
      fill = "steelblue",
      color = "white"
    ) +
    ggplot2::labs(
      title = "Distribution of Person-Level Episode Count Differences",
      x = "Algorithm count - Ground truth count",
      y = "Number of persons"
    ) +
    ggplot2::theme_minimal()
}
```

# Summary and key findings
```{r summary-key-findings}
# =========================================================
# Summary and key findings
# Includes a podium-oriented subset
# =========================================================

summary_df <- tibble::tibble(
  Metric = c(
    "Algorithm episodes",
    "Ground truth episodes",
    "Primary matched episodes (overlap)",
    "Algorithm match rate (primary overlap)",
    "Ground truth match rate (primary overlap)",
    "Overall classification accuracy",
    "Mean GA difference (days)",
    "GA within ±7 days",
    "GA within ±14 days",
    "Person exact count match",
    "Person count within ±1 episode"
  ),
  Value = c(
    match_results$primary_overlap$stats$total_algo,
    match_results$primary_overlap$stats$total_gt,
    match_results$primary_overlap$stats$matched,
    round(100 * match_results$primary_overlap$stats$match_rate_algo, 1),
    round(100 * match_results$primary_overlap$stats$match_rate_gt, 1),
    round(100 * class_metrics$overall_accuracy, 1),
    ifelse(nrow(ga_metrics$summary) > 0, round(ga_metrics$summary$mean_diff_days, 1), NA_real_),
    ifelse(nrow(ga_metrics$summary) > 0, round(100 * ga_metrics$summary$within_7_days, 1), NA_real_),
    ifelse(nrow(ga_metrics$summary) > 0, round(100 * ga_metrics$summary$within_14_days, 1), NA_real_),
    round(100 * person_metrics$summary$exact_match_rate, 1),
    round(100 * person_metrics$summary$within_1_rate, 1)
  )
)

summary_df %>%
  knitr::kable(caption = "Overall Validation Summary") %>%
  kableExtra::kable_styling(
    bootstrap_options = c("striped", "hover"),
    full_width = FALSE
  )
```



## Podium abstract subset:
```{r podium-abstract}
# Podium-oriented subset:
# emphasize episode identification and gestational-age agreement

podium_df <- tibble::tibble(
  Metric = c(
    "Primary matched episodes (overlap)",
    "Algorithm match rate (primary overlap)",
    "Ground truth match rate (primary overlap)",
    "Mean GA difference (days)",
    "GA within ±7 days",
    "GA within ±14 days"
  ),
  Value = c(
    match_results$primary_overlap$stats$matched,
    round(100 * match_results$primary_overlap$stats$match_rate_algo, 1),
    round(100 * match_results$primary_overlap$stats$match_rate_gt, 1),
    ifelse(nrow(ga_metrics$summary) > 0, round(ga_metrics$summary$mean_diff_days, 1), NA_real_),
    ifelse(nrow(ga_metrics$summary) > 0, round(100 * ga_metrics$summary$within_7_days, 1), NA_real_),
    ifelse(nrow(ga_metrics$summary) > 0, round(100 * ga_metrics$summary$within_14_days, 1), NA_real_)
  )
)

podium_df %>%
  knitr::kable(caption = "Podium-Focused Result Subset") %>%
  kableExtra::kable_styling(
    bootstrap_options = c("striped", "hover"),
    full_width = FALSE
  )
```


## Key findings
```{r key-findings}
cat("Key Findings:\n")
cat(
  "1. Primary overlap-based episode matching identified",
  match_results$primary_overlap$stats$matched,
  "matched episodes, corresponding to an algorithm-side match rate of",
  round(100 * match_results$primary_overlap$stats$match_rate_algo, 1),
  "% and a ground-truth-side match rate of",
  round(100 * match_results$primary_overlap$stats$match_rate_gt, 1),
  "%.\n"
)

cat(
  "2. Gestational age agreement among primary matched episodes was",
  ifelse(nrow(ga_metrics$summary) > 0, paste0(round(100 * ga_metrics$summary$within_14_days, 1), "% within ±14 days"), "not available"),
  ".\n"
)

cat(
  "3. Outcome classification accuracy remained secondary, with overall accuracy of",
  round(100 * class_metrics$overall_accuracy, 1),
  "% among primary matched episodes.\n"
)

cat(
  "4. At the person level, exact episode-count agreement was",
  round(100 * person_metrics$summary$exact_match_rate, 1),
  "%, increasing to",
  round(100 * person_metrics$summary$within_1_rate, 1),
  "% when allowing ±1 episode.\n"
)

cat(
  "5. Sensitivity analyses across end-date windows (±14 / ±30 / ±60 days) are reported to contextualize the primary overlap-based analysis.\n"
)
```






