---
title: "OMOP Pregnancy pipeline on Databricks"
format:
  html:
    toc: true
    code-fold: false
execute:
  echo: true
  warning: false
  message: true
params:
  # Path of the checked-out repository (Databricks Repo / Git folder)
  repo_path: "/Workspace/Repos/<your user>/OMOP-Pregnancy"
  # catalog.schema of the CDM tables and of the vocabulary (concept table)
  cdm_schema: "omop.data"
  vocabulary_schema: "omop.vocabulary"
  # catalog.schema for the pregnancy_episodes table; "" skips the table
  results_schema: ""
  # folder for CSV/RDS output; "" uses <repo>/output (git-ignored)
  output_folder: ""
  min_age: 15
  max_age: 56
---

This document runs the pregnancy episode pipeline on a Databricks cluster's
Spark session: no JDBC driver, no token. Attach the notebook to an
all-purpose cluster running a Databricks Runtime (not a SQL warehouse); the
runtime provides R, sparklyr and Java. The same steps are in
`inst/scripts/run_in_databricks.R` as a single script, and in
`Databricks/run_pipeline_notebook.r`, which the Repo opens as a notebook.

Fill in the parameters above, then run the chunks in order.

## 1. Packages

Installs anything missing. `SqlRender` needs `rJava`, which builds against
the runtime's Java; if that step fails, install `rJava` on its own first.
`DatabaseConnector` is not needed on this path.

```{r packages}
needed <- c("sparklyr", "DBI", "SqlRender", "dplyr", "lubridate", "readr")
missing <- needed[!vapply(needed, requireNamespace, logical(1), quietly = TRUE)]
if (length(missing) > 0) {
  message("Installing: ", paste(missing, collapse = ", "))
  install.packages(missing)
}
invisible(lapply(needed, requireNamespace, quietly = TRUE))
suppressPackageStartupMessages({
  library(dplyr)
  library(lubridate)
  library(readr)
})
cat("R", R.version.string, "\n")
cat("sparklyr", as.character(packageVersion("sparklyr")), "\n")
cat("SqlRender", as.character(packageVersion("SqlRender")), "\n")
```

## 2. Configuration

```{r configuration}
repo_path         <- params$repo_path
cdm_schema        <- params$cdm_schema
vocabulary_schema <- if (nzchar(params$vocabulary_schema)) params$vocabulary_schema else cdm_schema
results_schema    <- if (nzchar(params$results_schema)) params$results_schema else NULL
output_folder     <- if (nzchar(params$output_folder)) params$output_folder else file.path(repo_path, "output")
min_age           <- params$min_age
max_age           <- params$max_age

stopifnot(dir.exists(repo_path))
setwd(repo_path)
if (!dir.exists(output_folder)) dir.create(output_folder, recursive = TRUE)

cat("Repository:        ", repo_path, "\n")
cat("CDM schema:        ", cdm_schema, "\n")
cat("Vocabulary schema: ", vocabulary_schema, "\n")
cat("Results schema:    ", if (is.null(results_schema)) "(none)" else results_schema, "\n")
cat("Output folder:     ", output_folder, "\n")
```

## 3. Load the pipeline code

The package is used from source; nothing is installed into the R library.

```{r load-code}
for (dir in c("R/00_connection", "R/00_concepts", "R/01_extraction",
              "R/02_algorithms", "R/03_results", "R/03_utilities")) {
  for (f in list.files(dir, pattern = "\\.R$", full.names = TRUE)) source(f)
}
source("R/main.R")
cat("Loaded pipeline code from", repo_path, "\n")
```

## 4. Connect to the Spark session

```{r connect}
connection <- create_spark_connection(
  cdm_schema = cdm_schema,
  vocabulary_schema = vocabulary_schema,
  results_schema = results_schema
)
print(connection)

# Quick check that the CDM is reachable
n_persons <- db_query(connection, paste0("SELECT COUNT(*) AS n FROM ", cdm_schema, ".person"))
cat("Persons in CDM:", format(n_persons$n[1], big.mark = ","), "\n")
```

If `create_spark_connection()` cannot attach to the session on your runtime,
open it yourself and pass it in:

```{r connect-manual, eval=FALSE}
sc <- sparklyr::spark_connect(method = "databricks")
connection <- create_spark_connection(cdm_schema, vocabulary_schema, results_schema, sc = sc)
```

## 5. Concept sets

```{r concepts}
concepts <- load_concept_sets()
cat("HIP concepts:", nrow(concepts$hip_concepts), "\n")
cat("PPS concepts:", nrow(concepts$pps_concepts), "\n")
cat("Outcome categories:", nrow(concepts$matcho_limits), "\n")
```

## 6. Extract records

The only step that touches the database. Person cohort, HIP and PPS
concept lists and the ESD concept list become temporary views in the
session; every record they match is brought into R.

```{r extract}
start_time <- Sys.time()

cohort_data <- extract_pregnancy_cohort(
  connection = connection,
  cdm_schema = cdm_schema,
  vocabulary_schema = vocabulary_schema,
  hip_concepts = concepts$hip_concepts,
  pps_concepts = concepts$pps_concepts,
  min_age = min_age,
  max_age = max_age
)

cat(sprintf("Extraction took %.1f minutes\n",
            as.numeric(difftime(Sys.time(), start_time, units = "mins"))))
cat("Persons:            ", n_distinct(cohort_data$persons$person_id), "\n")
cat("Condition records:  ", nrow(cohort_data$conditions), "\n")
cat("Procedure records:  ", nrow(cohort_data$procedures), "\n")
cat("Observation records:", nrow(cohort_data$observations), "\n")
cat("Measurement records:", nrow(cohort_data$measurements), "\n")
cat("PPS timing records: ", nrow(cohort_data$gestational_timing), "\n")
cat("ESD timing records: ", nrow(cohort_data$esd_timing), "\n")
```

## 7. HIP: outcome-based episodes

```{r hip}
hip_episodes <- run_hip_algorithm(cohort_data, concepts$matcho_limits, concepts$matcho_outcome_limits)
cat("HIP episodes:", nrow(hip_episodes), "\n")
hip_episodes %>% count(outcome_category, sort = TRUE)
```

## 8. PPS: timing-based episodes

```{r pps}
pps_episodes <- run_pps_algorithm(cohort_data, concepts$pps_concepts)
cat("PPS episodes:", nrow(pps_episodes), "\n")
pps_episodes %>% count(outcome_category, sort = TRUE)
```

## 9. Merge

```{r merge}
merged_episodes <- merge_pregnancy_episodes(hip_episodes, pps_episodes)
cat("Merged episodes:", nrow(merged_episodes), "\n")
merged_episodes %>% count(algorithm_used, sort = TRUE)
```

## 10. Estimated start dates and quality flags

```{r esd}
final_episodes <- calculate_estimated_start_dates(merged_episodes, cohort_data, concepts$pps_concepts)
final_episodes <- add_episode_quality_metadata(final_episodes, concepts$matcho_limits)

cat("Episodes:", nrow(final_episodes), "for", n_distinct(final_episodes$person_id), "persons\n")
final_episodes %>% count(outcome_category, sort = TRUE)
final_episodes %>% count(precision_category, sort = TRUE)
final_episodes %>% count(term_duration_flag, outcome_concordance_score)
```

## 11. Save

CSV, RDS and a summary go to the output folder. When a results schema is
set, the `pregnancy_episodes` table there is replaced.

```{r save}
save_results(
  episodes = final_episodes,
  output_folder = output_folder,
  connection = connection,
  results_schema = results_schema,
  start_time = start_time
)
list.files(output_folder)
cat(sprintf("Total runtime %.1f minutes\n",
            as.numeric(difftime(Sys.time(), start_time, units = "mins"))))
```

## 12. A look at the output

```{r inspect}
final_episodes %>%
  select(person_id, episode_number, outcome_category, episode_start_date, episode_end_date,
         gestational_age_days, precision_category, algorithm_used) %>%
  head(20)
```

## 13. Diagnostic: where do pregnancy-loss outcomes go?
The registry holds tens of thousands of SA/AB/ECT pregnancies and the CDM has the codes, but the
final output has very few. This cell re-runs the HIP stages one at a time and counts SA, AB and ECT
at each step so the stage that drops them is visible. Uses `cohort_data` and `hip_episodes` from above.

```{r loss-trace}
loss_cats <- c("SA", "AB", "ECT", "SB", "LB", "DELIV")
count_by_cat <- function(df, col = "outcome_category", label) {
  if (is.null(df) || nrow(df) == 0) { cat(sprintf("%-42s (empty)\n", label)); return(invisible(NULL)) }
  tab <- table(factor(df[[col]], levels = loss_cats))
  cat(sprintf("%-42s %s\n", label, paste(sprintf("%s=%d", names(tab), as.integer(tab)), collapse = "  ")))
}

all_records <- bind_rows(cohort_data$conditions, cohort_data$procedures,
                         cohort_data$observations, cohort_data$measurements) %>%
  filter(!is.na(event_date)) %>% arrange(person_id, event_date)
initial_cohort <- all_records %>% filter(category %in% c("LB", "SB", "DELIV", "ECT", "AB", "SA", "PREG"))

cat("Raw HIP concept records by category:\n")
count_by_cat(initial_cohort, "category", "  records")
cat(sprintf("%-42s %s\n", "  distinct person-days",
            paste(sprintf("%s=%d", loss_cats, sapply(loss_cats, function(k)
              nrow(distinct(filter(initial_cohort, category == k), person_id, event_date)))), collapse = "  ")))

lim <- concepts$matcho_outcome_limits
lb  <- process_outcome_category(initial_cohort, "LB", lim);  count_by_cat(lb,  label = "1 per-category: LB")
sb  <- process_outcome_category(initial_cohort, "SB", lim);  count_by_cat(sb,  label = "1 per-category: SB")
ect <- process_outcome_category(initial_cohort, "ECT", lim); count_by_cat(ect, label = "1 per-category: ECT")
ab  <- process_outcome_category(initial_cohort, c("AB", "SA"), lim); count_by_cat(ab, label = "1 per-category: AB+SA")
dl  <- process_outcome_category(initial_cohort, "DELIV", lim); count_by_cat(dl, label = "1 per-category: DELIV")

s2 <- add_stillbirth_episodes(lb, sb, lim);   count_by_cat(s2, label = "2 after add_stillbirth")
s3 <- add_ectopic_episodes(s2, ect, lim);     count_by_cat(s3, label = "3 after add_ectopic")
s4 <- add_abortion_episodes(s3, ab, lim);     count_by_cat(s4, label = "4 after add_abortion")
s5 <- add_delivery_episodes(s4, dl, lim);     count_by_cat(s5, label = "5 after add_delivery")

s5 <- s5 %>% arrange(person_id, outcome_date, outcome_category) %>% group_by(person_id) %>%
  mutate(episode_number = row_number()) %>% ungroup()
s6 <- add_gestational_age_info(s5, all_records, concepts$matcho_limits); count_by_cat(s6, label = "6 after gestation matching")
s7 <- calculate_hip_start_dates(s6, concepts$matcho_limits)
s8 <- validate_hip_episodes(s7);              count_by_cat(s8, label = "7 after validation (final outcome)")
cat("   reclassified to PREG, by original category:\n")
print(s8 %>% filter(removed_outcome == 1) %>% count(removed_category, sort = TRUE))

count_by_cat(hip_episodes, label = "HIP output (run_hip_algorithm)")
count_by_cat(pps_episodes, label = "PPS output")
count_by_cat(merged_episodes, "HIP_outcome_category", "merge: HIP_outcome_category")
count_by_cat(merged_episodes, "PPS_outcome_category", "merge: PPS_outcome_category")
count_by_cat(merged_episodes, label = "merge: final outcome_category")
count_by_cat(final_episodes, label = "final (after ESD + metadata)")

cat("\nLoss episodes in HIP output whose merged outcome changed:\n")
merged_episodes %>%
  filter(HIP_outcome_category %in% c("SA", "AB", "ECT"), outcome_category != HIP_outcome_category) %>%
  count(HIP_outcome_category, PPS_outcome_category, outcome_category, algorithm_used, sort = TRUE) %>%
  print(n = 30)
```

The evaluation against the obstetric registry is a separate document,
`Evaluation/validation_report.qmd`, run with `gt_source: edw_databricks`.
